
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose the feedback arm of a valve positioner (the linkage connecting the valve's sliding stem to the positioner mechanism, telling the positioner how far open the valve is) comes loose, leaving the positioner ``thinking'' the control valve is always 50\% open even when it is not.

\vskip 10pt

Determine how this control valve will react when it receives a control signal starting at 4 mA and slowly climbing upward to 20 mA, assuming the positioner's calibration is fully closed at 4 mA and fully open at 20 mA.  Be as specific as you can in your answer.

\vskip 10pt

Assume that the control valve is Air-to-Open (Fail Closed).

\vfil

\underbar{file i00061}
\eject
%(END_QUESTION)





%(BEGIN_ANSWER)

This is a graded question -- no answers or hints given!

%(END_ANSWER)





%(BEGIN_NOTES)

At any signal value less than 12 mA, the positioner will drive the valve fully closed.  At any signal value greater than 12 mA, the positioner will drive the valve fully open.  With the feedback arm stuck at 50\%, the positioner essentially behaves as an ``on-off'' device, driving the valve open and shut any time the control signal falls below or rises above 12 mA.  This is similar to how an op-amp that's lost its negative feedback will behave as a comparator.

\vskip 10pt

The key to figuring this out is remembering that the job of any valve positioner is to drive as much or as little air pressure as necessary to the valve's actuator to get that valve to move to the desired position.  If the positioner sees a ``frozen'' control valve, it will absolutely saturate its output pressure in a vain effort to move that stuck valve.  In this case, the positioner will saturate low if the control signal is less than 50\% (because it sees the valve too far open, and tries as hard as it can to close it) and it will saturate high if the control signal is greater than 50\% (because it sees the valve too far close, and tries as hard as it can to open it).

%INDEX% Final Control Elements, valve: positioner (troubleshooting) 

%(END_NOTES)




