
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Locate the ``{\bf grade template}'' spreadsheet, and explore how you may calculate your own grade for any given second-year course by plugging in scores for exams (both mastery and proportional -- enter a number greater than 1 to simulate re-takes on a mastery exam), quizzes, tardiness, attendance, and other variables.  Note the difference between the grading for this INST200 course and any of the higher-credit courses (e.g. INST240, INST250, INST260). 

\vskip 10pt

Discuss the importance of signing a FERPA form, to allow prospective employers access to the details of your academic performance.  Without signing this release form, an instructor cannot comment at all about a student except to reveal whether or not they are enrolled in a course!

\vskip 10pt

Examine the grade spreadsheet to identify data an employer might be interested in seeing when considering a student for hire.  {\it Note that there is much more recorded here than just the final letter grade or percentage score!}  Which piece of data in these records do you suppose most employers focus on primarily?

\vskip 20pt

Also locate the quarter {\bf calendar} spreadsheet, and identify how to map calendar dates to the days shown on the front page of each section worksheet.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Why do you suppose the instructor provides you with a spreadsheet to calculate your own grade, rather than simply tell you what grade(s) you've earned throughout the time you spend in this program?
\item{} Why do you suppose there is a separate calendar document mapping worksheet sections and days to days of the year?  Why not just have those days of the year embedded in the PDF worksheet documents (on page 1)?
\end{itemize}

\underbar{file i00118}
%(END_QUESTION)





%(BEGIN_ANSWER)

Hint: both the grade template and the calendar spreadsheet files are found on the {\tt Y:} network drive at BTC, in Microsoft Excel format.

%(END_ANSWER)





%(BEGIN_NOTES)


%INDEX% Course organization, grading: spreadsheet-based grade calculator
%INDEX% Course organization, schedule: spreadsheet-based calendar

%(END_NOTES)


