
%(BEGIN_QUESTION)
% Copyright 2011, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read selected portions of the {\it IEC 62591 WirelessHART System Engineering Guide} (revision 2.2) and answer the following questions:

\vskip 10pt

Describe the ``{\bf Rule of 5 minimum}'' articulated in this document and give one realistic example of a {\sl Wireless}HART network not in compliance with the rule.

\vskip 10pt

Describe the ``{\bf Rule of 3}'' articulated in this document and give one realistic example of a {\sl Wireless}HART network not in compliance with the rule.

\vskip 10pt

Describe the ``{\bf Rule of 25\%}'' articulated in this document and give one realistic example of a {\sl Wireless}HART network not in compliance with the rule.

\vskip 10pt

Describe the ``{\bf Rule of 1}'' articulated in this document (for {\it control} applications only) and give one realistic example of a {\sl Wireless}HART network not in compliance with the rule.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain the rationale for each one of these rules!
\item{} Suppose the reported RSSI (Received Signal Strength Indication) value for a {\sl Wireless}HART device is $-40$ dBm.  Convert this decibel figure into {\it milliwatts} without using an electronic calculator.
\item{} Suppose the reported RSSI (Received Signal Strength Indication) value for a {\sl Wireless}HART device is $-53$ dBm.  Convert this decibel figure into {\it milliwatts} without using an electronic calculator.
\end{itemize}

\underbar{file i00542}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

{\bf Rule of 5 minimum:} there should be at least 5 field devices within direct range of the gateway in order to form a robust wireless mesh network.  A {\sl Wireless}HART network will function with fewer than 5 devices within direct range of the gateway, but the mesh network's foundation will not be as strong.  (Page 39 of 82)

\vskip 10pt

{\bf Rule of 3:} there should be at least 3 neighboring field devices within direct range of each field device, in order to form a robust mesh network with all field devices.  This ensures signal path redundancy.  (Page 39 of 82) 

\vskip 10pt

{\bf Rule of 25\%:} at least 25\% of the field devices in any network containing more than 5 devices should be within direct range of the gateway.  10\% is the absolute bare-minimum, but 25\% is more practical.  (Pages 40-43 of 82)

\vskip 10pt

{\bf Rule of 1:} every control (actuator) field device should be within direct range of the gateway for minimum latency.  (Page 46 of 82) 


\vskip 30pt

\noindent
$-40$ dBm is $-40$ dB less than 1 milliwatt.  Since each $-10$ dB is a reduction of one-tenth, $-40$ dB is one-tenth times one-tenth times one-tenth times one-tenth, which is $1 \over 10000$.  Therefore, $-40$ dBm is equivalent to a power value of $1 \times 10^{-4}$ milliwatts, or $1 \times 10^{-7}$ watts.

\vskip 10pt

\noindent
$-53$ dBm is $-53$ dB less than 1 milliwatt.  Since each $-10$ dB is a reduction of one-tenth and each $-3$ dB is a reduction of one-half, $-53$ dB is one-tenth times one-tenth times one-tenth times one-tenth times one-tenth times one-half, which is $1 \over {2 \times 10^5}$ or $5 \times 10^{-6}$.  Therefore, $-53$ dBm is equivalent to $5 \times 10^{-6}$ milliwatts, or $5 \times 10^{-9}$ watts (5 nW).  


% ADD: show an overhead site plan for some installation and ask students to assess whether any of these rules are broken by the layout.









\vfil \eject

\noindent
{\bf Summary Quiz:}

Identify the rationale for the ``Rule of 1'' when using {\sl Wireless}HART for closed-loop process control applications.

\begin{itemize}
\item{} Provide redundancy in case the gateway fails
\vskip 5pt 
\item{} Prevent interference from other devices
\vskip 5pt 
\item{} Minimize latency time (transit time delay) 
\vskip 5pt 
\item{} Reduce battery power consumption
\vskip 5pt 
\item{} Ensure all devices are on the same gateway
\vskip 5pt 
\item{} Allows faster acquisition of new devices
\end{itemize}


%INDEX% Reading assignment: IEC 62591 WirelessHART System Engineering Guide

%(END_NOTES)


