
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

%Magnetic flowmeters exhibit special advantages and disadvantages when compared to other flow-measuring technologies.  For each of the following strengths and weaknesses, explain {\it why} it is this way for a magnetic flowmeter:
Elektromagnetiske flowmeter har spesielle fordeler og ulemper sammenlignet med andre flowm{\aa}lere. For hver av fordel og ulempe forklar hvorfor. 
\vskip 10pt

{\bf Fordeler:}

\medskip
\item{$\bullet$} Ikke krav til lange rettstrekk opp- og nedstr{\o}ms. Typisk 5DN opp og 3DN ned. %Short upstream/downstream straight-pipe requirements: 5 up and 3 down (typically)
\item{$\bullet$} Utgangen er linear med volumetiske str{\o}mningsrate. (ikke behov til kvardratrot uttekning) %Output is linearly related to volumetric flow rate -- no square root characterization required
\item{$\bullet$} Good rangeability
\item{$\bullet$} Bidireksjonal m{\aa}ling er mulig. %Bidirectional measurement possible
\medskip

\vskip 10pt

{\bf Ulemper:}

\medskip
\item{$\bullet$} Virker bare med ledende v{\ae}sker. %Does not work with nonconducting fluids
\item{$\bullet$} God jording er essensielt %Excellent electrical grounding of the flowmeter is {\it essential}
\item{$\bullet$} Lag p{\aa} elektrodene kan p{\aa}virke resultatet. %Coating of electrodes may affect performance
\item{$\bullet$} M{\aa} monteres horisontalt i r{\o}ret. (Aldig vertikalt) %Needs to be installed in pipe with electrodes horizontal, never vertical
\medskip

\vskip 10pt

%Suppose a magflow meter is operating with a partially-filled pipe, with both electrodes still fully contacting the liquid.  Will this operating condition cause a {\it zero shift}, a {\it span shift}, a {\it linearity error}, or a {\it hysteresis error}?  Explain your reasoning.
Anta at et elektromagnetiske flowmeter bare delvis er fylt opp, men begge elektrodene er dekket av v{\ae}ske. Vil dette for{\aa}rsake: nullpunktsforflyttingt, span forflyttning, linearitetsfeil eller hysteresefeil. Forklar hvordan du tenker.

\vskip 10pt

%Suppose the flowstream through a magflow meter contains some non-conductive solids in addition to conductive liquid.  Will this affect the accuracy or reliability of the flowmeter?  Explain why or why not.
Anta at mediet som str{\o}mmer igjennom et elektromagnetisk flowmeter inneholder ikke ledende faste klumper. Vil dette p{\aa}virke m{\aa}leresultatet? 

\vskip 10pt

\underbar{file i00525}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

{\bf Strengths:}

\medskip
\item{$\bullet$} Short upstream/downstream straight-pipe requirements: 5 up and 3 down -- {\it induced EMF is a good representation of average flow rate; directions of flow not perpendicular to magnetic field and to electrodes don't contribute much to the EMF}
\item{$\bullet$} Output is linearly related to volumetric flow rate -- no square root characterization required -- {\it electromagnetic induction is a linear function of velocity}
\item{$\bullet$} Good rangeability -- {\it linearity makes this possible; limited by ambient electrical noise, though}
\item{$\bullet$} Bidirectional measurement possible -- {\it polarity (phase) simply reverses}
\medskip

\vskip 10pt

{\bf Weaknesses:}

\medskip
\item{$\bullet$} Does not work with nonconducting fluids -- {\it induction requires a conductor}
\item{$\bullet$} Excellent electrical grounding of the flowmeter is {\it essential} -- {\it to minimize effects of electrical noise on measurement}
\item{$\bullet$} Coating of electrodes may affect performance -- {\it adds electrical resistance to measurement circuit}
\item{$\bullet$} Needs to be installed in pipe with electrodes horizontal, never vertical -- {\it so air bubble will have a harder time breaking the measurement circuit}
\medskip

\vskip 10pt

A partially-filled pipe will cause the velocity to be greater for any given volumetric flow rate (less effective cross-sectional area).  This multiplies the flowmeter's indication, causing it to exhibit a positive span error.

\vskip 10pt

The presence of non-conductive matter in the flowstream is of no consequence, so long as the electrodes do not become fouled with a non-conductive coating.  Otherwise, the presence of conductive liquid around the non-conductive matter will still ensure conductivity across the diameter of the magnetic flowmeter, and thus will continue to ensure its proper operation.

Students commonly think that the presence of non-conducting material in an otherwise conductive stream of liquid is somehow a deal-breaker for magflow meters.  Practical examples to the contrary include wastewater treatment and wood pulp flow measurement.


%INDEX% Measurement, flow: magnetic

%(END_NOTES)


