
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Why is pH measurement important in industry?  Who cares what the pH value of a solution is, anyway?  Identify some applications where we absolutely {\it do} care, and explain why.

\underbar{file i00617}
%(END_QUESTION)





%(BEGIN_ANSWER)

pH measurement and control is important for wastewater treatment, because extreme pH values (either too acidic or too caustic) can adversely affect both flora and fauna in the water body that the treated wastewater is discharged to.   

\vskip 10pt

In biopharmaceutical production, pH measurement and control is absolutely critical.  Batches of cells will not grow, or grow properly, if the pH is not kept within tight bounds.

\vskip 10pt

Food production (especially brewing) is another process where pH measurement (and control) is critical.  Poor pH control leads to bad beer, and no one wants that.

%(END_ANSWER)





%(BEGIN_NOTES)

In medical instrumentation, blood pH measurement is an important analytical function.  A deviation as little as +/- 0.2 pH is serious, and a deviation of +/- 0.4 pH is usually fatal (according to a ``Chemical Insight'' note on page 803 of {\it Chemistry, 2nd Edition} by Stanley R. Radel and Marjorie H. Navidi).

%INDEX% Chemistry, pH: importance to different industrial processes

%(END_NOTES)


