
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose a gas turbine engine at the gas-fired power plant you are working at vents 47,000 SCFM of exhaust gases while operating at full power.  The measured NO$_{2}$ concentration of the engine's exhaust gas is 5 ppm (by volume) at full power.  Calculate both the number of moles of NO$_{2}$ emitted and the mass of NO$_{2}$ emitted by this engine per 24-hour day, assuming it runs at full power the whole time.

Assume ``standard'' conditions of 1 atmosphere and 68$^{o}$F for the ``SCFM'' gas flow quantity.  Be sure to show all your work!


\vfil

\underbar{file i00682}
\eject
%(END_QUESTION)





%(BEGIN_ANSWER)

This is a graded question -- no answers or hints given!

%(END_ANSWER)





%(BEGIN_NOTES)

We are given the total volumetric gas flow rate out of the engine's exhaust (47,000 SCFM) and also the volumetric concentration of NO$_{2}$ gas in this stream.  Simple multiplication will tell us how many SCFM of NO$_{2}$ gas we have:

$$\left( {47000 \hbox{ SCFM} \over 1} \right) \left( {5 \hbox{ parts NO}_2 \over 1000000 \hbox{ parts total}} \right) = 0.235 \hbox{ SCFM NO}_2 \hbox{ gas}$$

Over a period of 24 hours (1440 minutes), the accumulated total of NO$_{2}$ gas emitted by the engine will be 338.4 SCF.  We may calculate how many moles of NO$_{2}$ gas this represents if we use the Ideal Gas Law.  First, converting SCF into liters:

$$\left( {338.4 \hbox{ ft}^3 \over 1} \right) \left( {1728 \hbox{ in}^3 \over 1 \hbox{ ft}^3 } \right) \left( {3.7854 \hbox{ liters} \over 231 \hbox{ in}^3} \right) = 9582.39 \hbox{ liters}$$

Solving for $n$ using the Ideal Gas Law:

$$PV = nRT$$

$$n = {PV \over RT} = {(1 \hbox{ atm})(9582.39 \hbox{ l}) \over (0.0821 \hbox{ l}\cdot\hbox{atm/mol}\cdot\hbox{K})(293.15 \hbox{ K})} = 398.14 \hbox{ mol}$$

\vskip 10pt

Now the task remains to calculate the mass of this NO$_{2}$ gas emitted in a 24-hour period.  Since NO$_{2}$ has a molecular weight equal to 14 + 2(16) = 46 amu (i.e. 46 grams per mole), we may calculate the mass in grams by multiplying that molecular weight by the molar quantity:

\vskip 10pt

$$m = \left(46 \hbox{ g} \over \hbox{mol} \right) \left({398.14 \hbox{ mol} \over 1} \right) = 18,314.6 \hbox{ grams} = 18.3146 \hbox{ kg}$$

Thus, there will be 18.3 kg of NO$_{2}$ emitted by this turbine engine per 24-hour period.

%INDEX% Chemistry, spectroscopy
%INDEX% Process: gas turbine engine NOx emissions

%(END_NOTES)


