
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

The following list of questions are often asked of interviewees applying for computer programming positions at the Microsoft Corporation:

\begin{itemize}
\item{} How are M\&Ms made?
\vskip 5pt
\item{} If you had a clock with lots of moving mechanical parts, you took it apart piece by piece without keeping track of the method of how it was disassembled, then you put it back together and discovered that 3 important parts were not included; how would you go about reassembling the clock?
\vskip 5pt
\item{} If you had to learn a new computer language, how would you go about doing it? {\it Translating this question into ``instrumentation'' terms, if you had to learn how a new type of instrument worked, how would you go about doing it?}
\vskip 5pt
\item{} You have been assigned to design Bill Gates' bathroom.  Naturally, cost is not a consideration.  You may not speak to Bill.
\vskip 5pt
\item{} If Microsoft told you we were willing to invest \$5 million in a start up of your choice, what business would you start?  Why?
\vskip 5pt
\item{} How would you explain how to use Microsoft Excel to your grandma?  {\it Translating this question into ``instrumentation'' terms, how would you explain your chosen career (Instrument Technician) to a small child in such a way that they know what kinds of activities you do at work on a typical day?}
\vskip 5pt
\item{} Suppose you go home, enter your house/apartment, hit the light switch, and nothing happens -- no light floods the room.  What exactly, in order, are the steps you would take in determining what the problem was?
\vskip 5pt
\item{} Why is it that when you turn on the hot water in any hotel, for example, the hot water comes pouring out almost instantaneously?
\vskip 5pt
\item{} Explain a scenario for testing a salt shaker.
\vskip 5pt
\item{} Interviewer hands you a black pen and says nothing but ``This pen is red.''
\end{itemize}

Choose any one of these questions, and try answering it as best you can.  Also, identify the most difficult question on this list and explain why it is difficult for you to answer.

\vskip 50pt

\underbar{file i00747}
%(END_QUESTION)





%(BEGIN_ANSWER)

Discuss these with your classmates, with an eye toward practical answers as well as figuring out what the interviewer is trying to determine with each question.

%(END_ANSWER)





%(BEGIN_NOTES)

Here are some other (difficult) interview questions, not necessarily asked at Microsoft interviews.  This list is a good one to use during a ``group interview'' with your students, selecting individual students one at a time to answer questions randomly from the list to see how they answer, and then to have the class offer suggestions and critique the first student's answer:

\begin{itemize}
\item{} What specific strengths will you bring to this job?
\item{} What do you think it takes for a person to be successful in this field, and why?
\item{} What do you look for in a job and a company?
\item{} What did you enjoy least about your last job?
\item{} What did you enjoy least about school?
\item{} Describe a situation where you assumed responsibilities beyond your immediate job description.
\item{} Why have you chosen this specific field of work?
\item{} Where do you see yourself professionally 5 years from now?  10 years from now?
\item{} Describe your ``dream job'' and explain what makes it so.
\item{} What do you do when things are slow at work?
\item{} Rank the following in order of importance to you: {\it pay}, {\it working hours}, {\it job duties}, and {\it work environment}.
\item{} Describe what factors most motivate you.
\item{} Describe the most recent skill you have learned.
\item{} How do you like to be rewarded for a job well done?
\item{} If you could re-do your college education, what would you do differently?
\item{} Explain how you manage stress.
\item{} Describe the sorts of people you find difficult to work with, and explain why.
\item{} Describe the sorts of people you find easy to work with, and explain why.
\item{} Describe an instance where you received helpful criticism from someone else.
\item{} Describe a specific instance in your life where you worked as a team to overcome a daunting problem.
\item{} What will you do in your first 30 days of working here?
\item{} List some of the advantages of diversity in the workplace.
\item{} Explain how you budget time.
\item{} Identify a type of decision that you find easy to make.
\item{} Identify a type of decision that you find difficult to make.
\item{} Identify a skill you would like to improve in the next year.
\item{} Describe a time when you were forced to change your mind about something significant, and explain why you did.
\item{} Identify situations or factors on a job that cause you to experience the greatest stress.
\item{} Describe an incident where a plan of yours failed, and explain how you were able to recover from that failure.
\item{} How do you plan to stay current on new technology within your chosen profession?
\item{} Describe the characteristics of poor management.
\item{} Describe the best management you have ever worked for, and explain what characteristics made that management good.
\item{} Describe a situation where you made an extraordinary contribution to a job.
\item{} If there was one thing you could change about your last job, what would that be, and why?
\item{} Describe an experience where you learned from making a mistake.
\item{} Describe a situation where you helped to resolve an interpersonal conflict.
\item{} Describe a situation where you had to manage a crisis (non-interpersonal), and the rationale behind your decisions.
\item{} Describe a time when you were charged with fixing a problem created by someone else, and the steps you took to fix it.
\item{} What natural aptitudes do you have that makes you well suited for your chosen career?
\item{} What do you hope to gain from this job that you are not receiving at your present job, and why is this important to you?
\end{itemize}

%INDEX% Career, interviewing: Microsoft sample questions
%INDEX% Career, interviewing: difficult interview questions

%(END_NOTES)


