
%(BEGIN_QUESTION)
% Copyright 2011, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Examine this page of program code from a Siemens APOGEE system (written in Siemens' ``PPCL'' programming language), then answer the following questions:

\vskip 10pt

\hbox{ \vrule
\vbox{ \hrule \vskip 3pt
\hbox{ \hskip 3pt
\vbox{ \hsize=6.5in \raggedright

\noindent
{\tt 04600 C *****SWAMP COOLER CONTROL*****}

\noindent
{\tt 04610 C}

\noindent
{\tt 04620 C IF ANY WEST ROOM IS GREATER THAN 2 DEG ABOVE ITS SETPOINT, THEN TURN ON} 

\noindent
{\tt 04630 C THE SWAMP COOLER, ELSE SHUT OFF SWAMP COOLER}

\noindent
{\tt 04640 C}

\noindent
{\tt 04650 IF ("MV.GH.1W:ROOM TEMP" .GT. "MV.GH.1W.CLG.STPT" + 2) THEN ON ("MV.GH.1W.CLG")}

\noindent
{\tt 04652 IF ("MV.GH.1W:ROOM TEMP" .LE. "MV.GH.1W.CLG.STPT") THEN OFF ("MV.GH.1W.CLG")}

\noindent
{\tt 04660 IF ("MV.GH.2W:ROOM TEMP" .GT. "MV.GH.2W.CLG.STPT" + 2) THEN ON ("MV.GH.2W.CLG")}

\noindent
{\tt 04662 IF ("MV.GH.2W:ROOM TEMP" .LE. "MV.GH.2W.CLG.STPT") THEN OFF ("MV.GH.2W.CLG")}

\noindent
{\tt 04670 IF ("MV.GH.3W:ROOM TEMP" .GT. "MV.GH.3W.CLG.STPT" + 2) THEN ON ("MV.GH.3W.CLG")}

\noindent
{\tt 04672 IF ("MV.GH.3W:ROOM TEMP" .LE. "MV.GH.3W.CLG.STPT") THEN OFF ("MV.GH.3W.CLG")}

\noindent
{\tt 04680 C}

\noindent
{\tt 04690 IF ("MV.GH.1W.CLG".OR."MV.GH.2W.CLG".OR."MV.GH.3W.CLG") THEN ON ("MV.GH.3W:DO 6")}

\noindent
\hskip 32pt {\tt ELSE OFF ("MV.GH.3W:DO 6")}

}
\hskip 3pt}%
\vskip 5pt \hrule}%
\vrule}

\vskip 10pt

What is the significance of the letter {\tt C} preceding many of the lines of code in this program, particularly the signficance of lines 04610, 04640, and 04680?

\vskip 10pt

Identify the meanings of the following ``operators'' in Siemens PPCL code: {\tt .LE.} and {\tt .GT.}

\vskip 10pt

Explain how lines 04650 through 04690 control the ``swamp cooler''.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Do you think it would be easier or more difficult to program custom control algorithms in a text-based language like PPCL or in a graphic-based language such as function blocks?
\item{} Is this program an example of {\it differential gap} (``on/off'') control or {\it proportional} control?
\item{} Identify any ``adjustments'' that may be made in the program to alter its quality of control (i.e. how tightly it regulates room temperature to setpoint).
\end{itemize}

\underbar{file i00709}
%(END_QUESTION)





%(BEGIN_ANSWER)

Each line beginning with a letter ``C'' is a {\it comment}, placed there strictly for the benefit of any human being reading the program.  Comments are ignored by the controller as it executes the code.

\vskip 10pt

The {\tt .LE.} operator stands for ``Less Than Or Equal To'' while the {\tt .GT.} operator stands for ``Greater Than''.

%(END_ANSWER)





%(BEGIN_NOTES)

If any of three different room temperature sensors exceed the setpoint value by 2 degrees, discrete output {\tt DO 6} will activate.  That discrete output returns to an inactive state when all three room temperature sensors are at or below the setpoint value.

%INDEX% DDC, Siemens APOGEE system
%INDEX% DDC, Siemens PPCL code (APOGEE DDC control)

%(END_NOTES)


