
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read selected portions of the ``Fisher Type 1098-EGR and 1098H-EGR Pilot-Operated Regulators'' product bulletin (document 71.2:1098-EGR), and answer the following questions:

\vskip 10pt

Page 4 shows excellent illustrations of the regulator body and actuator assemblies, as well as internal views of the pilot mechanism.  Identify the type of valve trim used to throttle process fluid (stem-guided, cage-guided, port-guided), and also whether the valve body is {\it direct} or {\it reverse} acting.  Additionally, explain how the valve plug position may be monitored while the regulator is in service.

\vskip 10pt

The purpose of the {\it pilot} is to provide a {\it loading pressure} to the actuator of the 1098 pressure regulator.  Page 6 explains how the pilot provides this loading pressure to the main actuator.  Explain what this ``loading pressure'' does to the main regulator mechanism, and also how the regulator self-adjusts to changes in the downstream pipe pressure.  Also, identify the fluid used to actuate the regulator -- it is {\it not} instrument air!

\vskip 10pt

What type of {\it stem packing} do you see used in this valve?  Explain why you think the valve is designed in this way.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} A powerful problem-solving technique is performing a {\it thought experiment} where you mentally simulate the response of a system to some imagined set of conditions.  Describe a useful ``thought experiment'' for this system, and how the results of that thought experiment are helpful to answering the question.
\item{} Fisher's {\it Fundamentals of Gas Pressure Regulation} (Technical Monograph 27) by Floyd D. Jury gives a fairly comprehensive explanation of pressure regulators, including pilot-operated regulators.  No photographs in this document, but a good amount of simple illustrations are provided.  If you have lingering questions after reading the Fisher 1098 product bulletin, peruse this monograph.
\end{itemize}

\underbar{file i00783}
%(END_QUESTION)





%(BEGIN_ANSWER)

The stem packing question is a ``trick'': there is no stem packing at all in this valve!  The reason no packing is required is because any leakage past the stem will simply enter the diaphragm housing and then pass to the downstream side of the valve through the outlet pressure feedback tube, becoming part of the regulator's out-flow.  These regulators normally operate in throttling mode, and tight shut-off is not critical.

%(END_ANSWER)





%(BEGIN_NOTES)

A {\it regulator} is a special form of control valve that is directly actuated by the process.  In other words, a regulator is its own, self-contained control loop in a single device.  A normal control valve receives an actuation signal from a {\it controller}, which in turn receives a measured variable (process variable) signal from a {\it transmitter}, and as such is but a single component in a complete control {\it system}.

Examples of regulators include pressure regulators (both gas and liquid), flow regulators, level regulators, and temperature regulators.

\vskip 10pt

The Fisher 1098-EGR pressure regulators use {\it reverse-acting} valve trim (i.e. the valve opens up as the stem is pushed into the valve body!).  A travel indicator serves to show the position of the valve stem while the regulator is in operation.  These are {\it cage-guided} valves.

\vskip 10pt

The ``loading pressure'' supplied by the pilot regulator pre-loads the 1098-EGR's diaphragm with natural gas pressure to open the regulator's trim, allowing more natural gas to flow through and increase downstream pressure.  This downstream pressure in turn is sensed on the other side of the 1098-EGR's diaphragm to achieve a condition of force-balance with the spring and the loading pressure.

%INDEX% Final Control Elements, valve: regulator
%INDEX% Reading assignment: Fisher 1098EGR regulator bulletin

%(END_NOTES)


