
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Sometimes, pneumatically actuated valves are described as being either {\it air-to-open} or {\it air-to-close}.  In light of what you know about ``direct-acting'' and ``reverse-acting'' valve mechanisms and actuators, what combination(s) of valve type and actuator type is necessary to make an ``air-to-open'' valve?  What combination(s) of valve type and actuator type is necessary to make an ``air-to-close'' valve?

\underbar{file i00780}
%(END_QUESTION)





%(BEGIN_ANSWER)

For an ``air-to-open'' valve, either of these two combinations will work:

\begin{itemize}
\item{} Direct-acting valve + reverse-acting actuator
\item{} Reverse-acting valve + direct-acting actuator
\end{itemize}

\vskip 10pt

For an ``air-to-close'' valve, either of these two combinations will work:

\begin{itemize}
\item{} Direct-acting valve + direct-acting actuator
\item{} Reverse-acting valve + reverse-acting actuator
\end{itemize}

\vskip 10pt

It should be noted that direct-acting globe valves are far more common than reverse-acting globe valves.  In industry, you usually see direct acting valve bodies coupled to either direct- or reverse-acting actuators to configure the valve assemblies for air-to-close or air-to-open action, respectively.

%(END_ANSWER)





%(BEGIN_NOTES)

If a valve mechanism is direct-acting, it means it will open when the stem is pulled (upward) out of the body.  If attached to a reverse-acting actuator (stem ``pulls'' with air applied), the resulting valve will be air-to-open.  If attached to a direct-acting actuator (stem ``pushes'' outward with air applied), the resulting valve will be air-to-close.

If a valve mechanism is reverse-acting, it means it will close when the stem is pulled (upward) out of the body.  If attached to a reverse-acting actuator (stem ``pulls'' with air applied), the resulting valve will be air-to-close.  If attached to a direct-acting actuator (stem ``pushes'' outward with air applied), the resulting valve will be air-to-open.

%INDEX% Final Control Elements, valve: air-to-close
%INDEX% Final Control Elements, valve: air-to-open

%(END_NOTES)


