
%(BEGIN_QUESTION)
% Copyright 2014, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Sketch a diagram showing all wire connections between any model of Rosemount differential pressure transmitter, an Emerson DeltaV DCS controller, and a Fisher DVC6020 valve positioner to form a simple PID control system.  You may arbitrarily choose to use either ``M-series'' or ``S-series'' DeltaV I/O cards.  Your task is to locate the appropriate datasheets or manuals in the Instrumentation Reference (or online) to identify the proper wire terminals to connect, then sketch a simple loop diagram showing how the appropriate terminals on each device connect to terminals on the other devices to make a functional instrument loop.  You are free to choose which channel on the DeltaV I/O cards to use for your loop.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} A problem-solving technique useful for making proper connections in pictorial circuit diagrams is to first identify the directions of all DC currents entering and exiting component terminals, as well as the respective voltage polarity marks (+,$-$) for those terminals, based on your knowledge of each component acting either as an electrical {\it source} or an electrical {\it load}.  Discuss and compare how these arrows and polarity marks simplify the task of properly connecting wires between components. 
\item{} Locate a wiring diagram for one of the {\it redundant} I/O card options available for the DeltaV system (either M-series or S-series) and reference that diagram to explain how redundant I/O cards connect to the same field instrument.
\end{itemize}

\underbar{file i00818}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Have students show you their loop diagrams along with the pages in the equipment manuals showing all appropriate screw terminals on the instruments and I/O cards.  Check to make sure they are properly annotating their diagrams with voltage polarities and current direction arrows to help them figure out how to make all the proper connections.











\vfil \eject

\noindent
{\bf Summary Quiz:}

(This makes a good summary quiz in itself -- students showing they have performed these configuration exercises)


%INDEX% Documentation, loop diagram
%INDEX% Pictorial circuit sketching (field instrument wiring to process controller)

%(END_NOTES)


