
%(BEGIN_QUESTION)
% Copyright 2014, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the introduction to the ``Concepts of Probability'' section of the ``Process Safety and Instrumentation'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i00828}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Reliability is the probability that a device or system will perform its intended function.  This can be further divided into ``dependability'' and ``security'', terms defined within the electrical power industry to specify what is meant by reliable operation with regard to protective devices such as fuses and circuit breakers.  

\begin{itemize}
\item{} {\bf Dependability}: The probability a protective system will shut off power when needed 
\item{} {\bf Security}: The probability a protective system will allow power to flow when there is no danger
\item{} {\bf Reliability}: A combination of dependability and security 
\end{itemize}

\noindent
Extending these electrical-specific definitions to general use:

\begin{itemize}
\item{} {\bf Dependability}: The probability a system will take action when called upon to take action
\item{} {\bf Security}: The probability a protective system will take no action when it's not supposed to take any action
\item{} {\bf Reliability}: A combination of dependability and security 
\end{itemize}








\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Describe your own example of ``dependability'' as defined in the textbook.
\item{} Describe your own example of ``security'' as defined in the textbook.
\item{} Define ``dependability'' in a general sense, not in a specific sense such as how the electric power industry uses the term.
\item{} Define ``security'' in a general sense, not in a specific sense such as how the electric power industry uses the term.
\end{itemize}


%INDEX% Reading assignment: Lessons In Industrial Instrumentation, concepts of probability (introduction)

%(END_NOTES)

