
%(BEGIN_QUESTION)
% Copyright 2014, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Search through Chapter 1 (``DeltaV System Overview'') of the ``Getting Started with your DeltaV Digital Automation System'' manual (document D800002X122, March 2006) to answer the following questions:

\vskip 10pt

Based on the simple network diagram shown on page 1-1, what type of digital network connects the PC workstation(s) to the DCS controller?

\vskip 10pt

What is the {\it System Identifier} in a DeltaV system, and what purpose does it serve?

\vskip 10pt

Several software applications find frequent use in the maintenance and operation of a DeltaV control system: {\it DeltaV Explorer}, {\it DeltaV Operate Run}, {\it DeltaV Operate Configure}, {\it Control Studio}, and {\it DeltaV Books Online}.  Identify what each of these software tools does.



\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} A hallmark of Distributed Control Systems (DCSs) is {\it hardware redundancy}.  Identify specific redundant features of the Emerson DeltaV DCS.
\item{} Access a DeltaV workstation PC and try opening each of the software applications listed, exploring the features of each.  {\it Do not ``download'' or ``save'' anything, which will alter the configuration of the DCS -- just explore and observe!}
\end{itemize}

\underbar{file i00811}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

The Emerson DeltaV DCS uses redundant Ethernet cabling to connect PC workstations to controllers.  The only clue that this is Ethernet is the diagram showing two ``hubs'' where network cabling plug into (page 1-1).  Redundant processors are also supported.

\vskip 10pt

The System Identifier is a hardware key used to authenticate the DeltaV network (page 1-2).  It plugs into either the PC's parallel printer port or one of the USB ports.

\vskip 10pt

{\it DeltaV Explorer} allows you to explore and edit the component configurations for a DeltaV system (page 1-9).  {\it DeltaV Operate Run} is the interface software used by operators to actually run the system (page 1-15).  {\it DeltaV Operate Configure} is another mode of this software whereby the graphic operator screens are built and configured (page 1-10).  {\it Control Studio} is used to build and configure the control strategies making up a control system; literally, this is the programming interface for the DeltaV system (page 1-8).  {\it DeltaV Books Online} is a set of digital reference manuals for the DeltaV system, containing this very ``Getting Started'' manual!







\vfil \eject

\noindent
{\bf Prep Quiz:}

The {\it Control Studio} application software which comes as part of an Emerson DeltaV DCS is typically used for which purpose:

\begin{itemize}
\item{} Building graphic ``picture'' screens for operators to use
\vskip 5pt 
\item{} Commissioning new controller nodes on the DeltaV network
\vskip 5pt 
\item{} Operators switching between Auto and Manual controller modes
\vskip 5pt 
\item{} Calibrating the process transmitter so it will read accurately
\vskip 5pt 
\item{} Calibrating the control valve so that it will stroke accurately
\vskip 5pt 
\item{} Configuring the function blocks comprising a control module
\vskip 5pt 
\item{} Calculating the installed cost of your DeltaV control system
\end{itemize}


%INDEX% Reading assignment: Emerson DeltaV "Getting Started" manual (Chapter 1)

%(END_NOTES)


