
%(BEGIN_QUESTION)
% Copyright 2015, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose a radio transmitting antenna radiates 120 kW of RF power into space, and that only 30 milliwatts of this power reaches a receiving antenna some distance away.  Calculate the path loss in decibels {\it without using a calculator}.  Note that an addition or subtraction of 10 dB is exactly equal to a 10-fold multiplication or division ratio (respectively), and that an addition or subtraction of 3 dB is approximately equal to a 2-fold multiplication or division ratio.

\vskip 10pt

$L_{path}$ = \underbar{\hskip 50pt} dB

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain what creates path loss.  Is this something that can be overcome or at least mitigated by applying the correct technology?  Why or why not?
\item{} Explain what {\it fade loss} is, what creates it, and how it differs from path loss.
\item{} Is path loss affected by the type of antenna used to transmit the RF energy?  Explain why or why not.
\item{} Is path loss affected by the type of antenna used to receive the RF energy?  Explain why or why not.
\item{} Is path loss affected by the frequency of the RF energy transmitted?
\end{itemize}

\underbar{file i00957}
%(END_QUESTION)





%(BEGIN_ANSWER)

The ratio of 30 milliwatts to 120 kilowatts may be expressed in fractional form and then reduced into a product of halves and tenths:

$${30 \hbox{ mW} \over 120 \hbox{ kW}} = {0.030 \over 120000} = {3 \over 12000000} = \left({3 \over 12} \times {1 \over 1000000}\right) = \left({1 \over 4} \times {1 \over 1000000}\right)$$

$${30 \hbox{ mW} \over 120 \hbox{ kW}} = \left({1 \over 4} \times {1 \over 1000000}\right) = \left({1 \over 2} \times {1 \over 2} \times {1 \over 10} \times {1 \over 10} \times {1 \over 10} \times {1 \over 10} \times {1 \over 10} \times {1 \over 10}\right)$$

$${30 \hbox{ mW} \over 120 \hbox{ kW}} \approx \left(-3 \hbox{ dB} - 3 \hbox{ dB} -10 \hbox{ dB} -10 \hbox{ dB} -10 \hbox{ dB} -10 \hbox{ dB} -10 \hbox{ dB} -10 \hbox{ dB} \right)$$

$$L_{path} = {30 \hbox{ mW} \over 120 \hbox{ kW}} \approx - \hbox{\bf 66 dB}$$

%(END_ANSWER)





%(BEGIN_NOTES)


%INDEX% Electronics review: decibel power calculations (mental math)

%(END_NOTES)


