
%(BEGIN_QUESTION)
% Copyright 2015, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

\vskip 10pt

Identify any area(s) of your study in which you would like to become stronger.  Examples include technical reading, electrical circuit analysis, solving particular types of problems, time management, and/or skills applied in the lab.  Cite specific examples if possible, and bring these to your instructor's attention so that together you may target them for improvement.  As a starting point, try consulting the list of topics on the first page of the worksheet for the upcoming mastery exam, as well as the ``General Values and Expectations'' list near the beginning of the worksheet identifying the habits and qualities necessary for success in this career.

\vskip 20pt

Next, identify practical strategies you will use to strengthen these areas.  Examples include focusing on specific types of problem-solving whenever those types appear in the homework, working through practice problems for a particular subject, and/or coordinating with your lab team to give you more practice on specific skills.

\vskip 20pt

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} One useful strategy is to maintain a {\it journal} of all you've learned in a course of study.  Explore ways you could take the work you're already doing to prepare for homework (daily discussions with your instructor) and turn this into a journal or even a weblog (``blog'') for your own reflection and eventual use as a portfolio to showcase your capabilities to employers.
\item{} Where exactly are the practice problem worksheets located on the {\it Socratic Instrumentation} website?
\item{} Peruse the ``feedback questions'' for this (and/or past) course sections to identify any questions related to areas you would like to strengthen.
\end{itemize}

\underbar{file i00999}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)







\vfil \eject

\noindent
{\bf Prep Quiz:}

List any \underbar{two} of the topics covered on the upcoming mastery exam, other than {\it instrument range calculations (Question \#4)}, {\it algebra (Question \#5)}, or {\it circuit fault analysis (Question \#6)}.  Your list may include any of the ``review'' topics on the exam as well as topics from this course.








\vfil \eject

\noindent
{\bf Prep Quiz:}

List some of the fundamental principles (e.g. basic laws of physics, chemistry, mathematics, logic, electric circuits, process control, etc.) you have applied in this course of study so far, as well as a brief description of how you used these principles to solve problems and/or understand new concepts.


%INDEX% Metacognitive log, prompting students to reflect on principles and learning strategies
%INDEX% Reading Apprenticeship technique, metacognitive log

%(END_NOTES)


