
%(BEGIN_QUESTION)
% Copyright 2013, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

One of the unique features of this program is the inclusion of {\it mastery exams}, where students must answer questions with 100\% accuracy in order to pass.  Conventional ``proportional'' exams allow students to pass if a certain minimum score is achieved.  The problem with this testing strategy is that students may not actually learn {\it all} the concepts they're supposed to, but may still pass the exam if they are strong enough in the other concepts covered in that assessment.  The purpose of mastery exams is to guarantee proficiency in {\it all} critical concepts and not just {\it some}.

\vskip 10pt

Your instructor will hand out copies of the mastery exam for the INST200 ``Introduction to Instrumentation'' course, covering several critical concepts of circuit analysis taught in the first year of the Instrumentation program.  Do your best to answer all the questions correctly.  If you get any incorrect on the first attempt, the instructor will mark which {\it sections} (not which {\it questions}) you missed and return it to you for one more attempt.  If a mastery exam is not passed by the second attempt, it counts as a failed exam.

\vskip 10pt

Mastery exams may be re-taken any number of times with no grade penalty.  The purpose is to give students the constructive feedback and practice that they need in order to master all the concepts represented on the exam.  Every mastery exam must be passed before the next scheduled exam is given in order to receive a passing grade for that course, a period of approximately 2 weeks.  If any student is not able to pass a mastery exam with 100\% accuracy by the deadline date, they will receive an ``F'' grade for that course, and must re-take the course again during some future quarter.

\vskip 10pt

The INST200 mastery exam is given for the purpose of exposing students to this unique type of assessment.  Failing to pass the INST200 mastery exam will not result in a failing grade for the INST200 course, but students should be warned that poor performance on this exam often marks trouble in future Instrumentation courses, since so much of the second year's material builds on what was taught during the first year.   %Re-takes of the INST200 mastery exam will be scheduled throughout this quarter.  This exam must be passed in its entirety in order to earn a passing grade for the INST200 course.

\vskip 10pt

\underbar{file i01230}
%(END_QUESTION)





%(BEGIN_ANSWER)

 
%(END_ANSWER)





%(BEGIN_NOTES)


%INDEX% Course organization, assessment: mastery (practice)
%INDEX% Mastery exam practice

%(END_NOTES)


