
%(BEGIN_QUESTION)
% Copyright 2013, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

The nameplate on an electric compressor provides the following data:

\begin{itemize}
\item{} 4160 V / 60 Hz
\item{} Three-phase
\item{} 850 HP
\item{} 1800 RPM
\item{} Max. Discharge Pressure = 84 PSIG
\item{} Max. Flow = 1280 SCFM
\end{itemize}

\vskip 10pt

\noindent
Answer the following questions based on this nameplate data:

\vskip 10pt

Identify whether this is an {\it induction} motor or a {\it synchronous} motor.  If it were the {\it other} type of motor, how would the nameplate data differ?

\vskip 10pt

Determine the number of {\it poles} in the stator of this motor.  Does it have 2 poles, 4 poles, 6 poles, or more poles?

\vskip 10pt

Calculate the {\it slip} speed of this motor.

\vskip 10pt

\underbar{file i01233}
%(END_QUESTION)





%(BEGIN_ANSWER)

This is most definitely a synchronous motor and not an inducation motor.  It it were an induction motor, the speed would be less than synchronous.
 
\vskip 10pt

The number of poles in this motor is 4.

\vskip 10pt

Slip speed = 0 RPM (because it is synchronous).

%(END_ANSWER)





%(BEGIN_NOTES)

Current = 3.39 amps AC ($0.5 \times 746 \over 110$)

\vskip 10pt

This is most definitely an induction motor and not a synchronous motor, because the rated speed (3400 RPM) is below synchronous (3600 RPM).
 
\vskip 10pt

This is a {\it 2-pole} motor, because its synchronous speed is 3600 RPM.  If it were 4-pole, it would have a synchronous speed of 1800 RPM; 6 pole would be 1200 RPM, etc.

\vskip 10pt

Slip speed = 3600 RPM - 3400 RPM = 200 RPM

%INDEX% Electronics review: AC motor horsepower calculation

%(END_NOTES)


