
%(BEGIN_QUESTION)
% Copyright 2012, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Examine this page of program code from a Siemens APOGEE system (written in Siemens' ``PPCL'' programming language), then answer the following questions:

\vskip 10pt

\hbox{ \vrule
\vbox{ \hrule \vskip 3pt
\hbox{ \hskip 3pt
\vbox{ \hsize=6.5in \raggedright

\noindent
{\tt 09390 C}

\noindent
{\tt 09400 C ROOM TEMP ALARMS}

\noindent
{\tt 09410 C}

\noindent
{\tt 09420 IF("MV.RM122:ROOM TEMP" .GT. "MV.RM122.RTHL" .OR.}

\noindent
{\tt \hskip 30pt "MV.RM122:ROOM TEMP" .LT. "MV.RM122.RTLL") THEN ON("MV.RM122.RTA")}

\noindent
{\tt 09430 IF("MV.RM122:ROOM TEMP" .LE. "MV.RM122.RTHL" - 2 .AND.}

\noindent
{\tt \hskip 30pt "MV.RM122:ROOM TEMP" .GE. "MV.RM122.RTLL" + 2) THEN OFF("MV.RM122.RTA")}

}
\hskip 3pt}%
\vskip 5pt \hrule}%
\vrule}

\vskip 10pt

Suppose the programmed parameters for room 122 are as follows: the room temperature low limit is set to 60 $^{o}$F and the room temperature high limit is set to 80 $^{o}$F.  Based on these parameter values and the code shown above, answer the following questions:

\vskip 10pt

\begin{itemize}
\item{} Will the room temperature alarm bit be set (1) when the room temperature {\it reaches} 80 $^{o}$F, or when it {\it exceeds} 80 $^{o}$F?
\vskip 10pt
\item{} Will the room temperature alarm bit be set (1) when the room temperature {\it reaches} 60 $^{o}$F, or when it {\it falls below} 60 $^{o}$F?
\vskip 10pt
\item{} Identify the condition(s) which must be met in order to reset (0) the room temperature alarm bit.
\end{itemize}

\underbar{file i01208}
%(END_QUESTION)





%(BEGIN_ANSWER)

3 points for each of the answers identifying how the bit will be set, and 4 points for a complete answer (mentioning BOTH conditions) identifying how the bit will be reset.  No partial credit for the last answer (i.e. it must contain both conditions, not just one)!

\begin{itemize}
\item{} The room temperature alarm bit will be set when the room temperature {\bf exceeds} 80 $^{o}$F
\vskip 10pt
\item{} The room temperature alarm bit will be set when the room temperature {\bf falls below} 60 $^{o}$F
\vskip 10pt
\item{} In order to reset the room temperature alarm bit, the temperature must be between 62 $^{o}$F and 78 $^{o}$F inclusive
\end{itemize}

%(END_ANSWER)





%(BEGIN_NOTES)

{\bf This question is intended for exams only and not worksheets!}

%(END_NOTES)


