
%(BEGIN_QUESTION)
% Copyright 2012, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

The following segment of program code from a Siemens APOGEE system (written in Siemens' ``PPCL'' programming language) overrides the position of an outside air damper based on measured concentration levels of carbon dioxide gas (CO$_{2}$) exhaled by people in that room.  Examine this code, and then answer the following questions:

\vskip 10pt

\hbox{ \vrule
\vbox{ \hrule \vskip 3pt
\hbox{ \hskip 3pt
\vbox{ \hsize=6.5in \raggedright

\noindent
{\tt 08230 C HVU-2 OUTSIDE AIR DAMPER POSITION OVERRIDDEN} 

\noindent
{\tt 08240 C IF SPACE CO2 LEVELS RISE MORE THAN 300 PPM ABOVE SETPOINT}

\noindent
{\tt 08250 C}

\noindent
{\tt 08260 SAMPLE(300) IF("MV.HVU2.STATUS" .EQ. ON .AND. "MV.HVU2.CO2" .GT. "MV.OACO2" + 300)}

\noindent
{\tt \hskip 30pt THEN ON("MV.HVU2.CO2.OVRD") ELSE OFF("MV.HVU2.CO2.OVRD")}

\noindent
{\tt 08270 SAMPLE(30) IF("MV.HVU2.CO2.OVRD" .EQ. ON .AND. "MV.HVU.2:AIRTEMP" .GE. 52)}

\noindent
{\tt \hskip 30pt THEN "MV.HVU2.MOA" = "MV.HVU2.MOA" + 0.2}

\noindent
{\tt 08280 SAMPLE(5) IF("MV.HVU2.CO2.OVRD" .EQ. OFF .OR. "MV.HVU.2:AIRTEMP" .LE. 48)}

\noindent
{\tt \hskip 30pt THEN "MV.HVU2.MOA" = "MV.HVU2.MOA" - 0.2}
}

\hskip 3pt}%
\vskip 5pt \hrule}%
\vrule}

\vskip 10pt

Identify the variable in this code referring to the measured CO$_{2}$ concentration in the room (i.e. the {\it process variable}).

\vskip 10pt

Identify the variable in this code referring to the desired CO$_{2}$ concentration in the room (i.e. the {\it setpoint}).

\vskip 10pt

Identify the variable in this code referring to the position of the outside air damper for the room (i.e. the {\it output}).

\vskip 10pt

What condition must be met in order to allow the outside air damper to open up, in addition to the CO$_{2}$ concentration being excessive?

\vskip 10pt

What condition will cause the outside air damper to close down, even if the CO$_{2}$ concentration is too high?

\underbar{file i01211}
%(END_QUESTION)





%(BEGIN_ANSWER)

2 points for each of the answers:

\begin{itemize}
\item{} Identify the variable in this code referring to the measured CO$_{2}$ concentration in the room (i.e. the {\it process variable}) = {\tt MV.HVU2.CO2}
\vskip 10pt
\item{} Identify the variable in this code referring to the desired CO$_{2}$ concentration in the room (i.e. the {\it setpoint}) = {\tt MV.OACO2}
\vskip 10pt
\item{} Identify the variable in this code referring to the position of the outside air damper for the room (i.e. the {\it output}) = {\tt MV.HVU2.MOA}
\vskip 10pt
\item{} What condition must be met in order to allow the outside air damper to open up, in addition to the CO$_{2}$ concentration being excessive? {\bf The room air temperature must be at least 52 degrees.}
\vskip 10pt
\item{} What condition will cause the outside air damper to close down, even if the CO$_{2}$ concentration is too high?  {\bf If the room air temperature is 48 degrees or less.}
\end{itemize}


%(END_ANSWER)





%(BEGIN_NOTES)

{\bf This question is intended for exams only and not worksheets!}

%(END_NOTES)


