
%(BEGIN_QUESTION)
% Copyright 2015, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

The nameplate on an electric water pump provides the following data:

\begin{itemize}
\item{} 110 V / 60 Hz
\item{} Single-phase
\item{} 1/2 HP
\item{} 3400 RPM
\item{} Max. Lift = 22 ft
\item{} Max. Head = 115 ft
\item{} Max. Flow = 12 Gal/min
\item{} 1" outlet
\end{itemize}

\vskip 10pt

\noindent
Answer the following questions based on this nameplate data:

\vskip 10pt

Calculate the amount of line current expected in the motor's power wiring.

\vskip 10pt

Identify whether this is an {\it induction} motor or a {\it synchronous} motor.  If it were the {\it other} type of motor, how would the nameplate data differ?

\vskip 10pt

Determine the number of {\it poles} in the stator of this motor.  Does it have 2 poles, 4 poles, 6 poles, or more poles?

\vskip 10pt

Calculate the {\it slip} speed of this motor.

\vskip 10pt

\underbar{file i01232}
%(END_QUESTION)





%(BEGIN_ANSWER)

Current = 3.39 amps AC

\vskip 10pt

This is most definitely an induction motor and not a synchronous motor!
 
%(END_ANSWER)





%(BEGIN_NOTES)

Current = 3.39 amps AC ($0.5 \times 746 \over 110$)

\vskip 10pt

This is most definitely an induction motor and not a synchronous motor, because the rated speed (3400 RPM) is below synchronous (3600 RPM).
 
\vskip 10pt

This is a {\it 2-pole} motor, because its synchronous speed is 3600 RPM.  If it were 4-pole, it would have a synchronous speed of 1800 RPM; 6 pole would be 1200 RPM, etc.

\vskip 10pt

Slip speed = 3600 RPM $-$ 3400 RPM = 200 RPM

%INDEX% Electronics review: AC motor horsepower calculation

%(END_NOTES)


