
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Control valve types (e.g. globe, ball, butterfly) may be given relative flow capacity ratings called the $C_d$ factor.  This is similar to the concept of flow capacity ($C_v$), but generalized to a specific type or design of control valve.  The equation relating $C_d$ to $C_v$ is as follows:

$$C_d = {C_v \over d^2}$$

\noindent
Where,

$C_d$ = Relative flow capacity of the valve type

$C_v$ = Maximum flow capacity of the particular valve

$d$ = Nominal pipe size for the particular valve, inches

\vskip 10pt

Several valve capacity factors ($C_d$) for different control valve types are shown here, assuming full-area trim and a full-open position:
 
% No blank lines allowed between lines of an \halign structure!
% I use comments (%) instead, so that TeX doesn't choke.

$$\vbox{\offinterlineskip
\halign{\strut
\vrule \quad\hfil # \ \hfil & 
\vrule \quad\hfil # \ \hfil \vrule \cr
\noalign{\hrule}
%
% First row
{\bf Valve design type} & $C_d$ \cr
%
\noalign{\hrule}
%
% Another row
Single-port globe valve, ported plug & 9.5 \cr
%
\noalign{\hrule}
%
% Another row
Single-port globe valve, contoured plug & 11 \cr
%
\noalign{\hrule}
%
% Another row
Single-port globe valve, characterized cage & 15 \cr
%
\noalign{\hrule}
%
% Another row
Double-port globe valve, ported plug & 12.5 \cr
%
\noalign{\hrule}
%
% Another row
Double-port globe valve, contoured plug & 13 \cr
%
\noalign{\hrule}
%
% Another row
Rotary ball valve, segmented & 25 \cr
%
\noalign{\hrule}
%
% Another row
Rotary ball valve, standard port (diameter $\approx$ 0.8$d$) & 30 \cr
%
\noalign{\hrule}
%
% Another row
Rotary butterfly valve, 60$^{o}$, no offset seat & 17.5 \cr
%
\noalign{\hrule}
%
% Another row
Rotary butterfly valve, 90$^{o}$, offset seat & 29 \cr
%
\noalign{\hrule}
%
% Another row
Rotary butterfly valve, 90$^{o}$, no offset seat & 40 \cr
%
\noalign{\hrule}
} % End of \halign 
}$$ % End of \vbox

This $C_d$ data for different control valve types allows us to approximate any valve's full-flow $C_v$ factor knowing the type of valve and the pipe size.  Calculate the full-flow $C_v$ values for these control valves:

\begin{itemize}
\item{} Segmented ball valve, 4 inch pipe size; $C_v$ = \underbar{\hskip 50pt}
\vskip 10pt 
\item{} Single-port, cage-guided globe valve, 6 inch pipe size; $C_v$ = \underbar{\hskip 50pt}
\vskip 10pt 
\item{} Double-port, ported-plug globe valve, 2 inch pipe size; $C_v$ = \underbar{\hskip 50pt}
\vskip 10pt 
\item{} 90$^{o}$ butterfly valve with offset seat, 20 inch pipe size; $C_v$ = \underbar{\hskip 50pt}
\end{itemize}

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Given a certain required $C_v$ rating, which design of control valve allows the smallest valve (pipe) size?
\item{} Given a certain required $C_v$ rating, which design of control valve requires the largest valve (pipe) size?
\end{itemize}

\underbar{file i01371}
%(END_QUESTION)





%(BEGIN_ANSWER)

\begin{itemize}
\item{} Segmented ball valve, 4 inch pipe size; $C_v$ = \underbar{\bf 400}
\vskip 5pt 
\item{} Single-port, cage-guided globe valve, 6 inch pipe size; $C_v$ = \underbar{\bf 540}
\vskip 5pt 
\item{} Double-port, ported-plug globe valve, 2 inch pipe size; $C_v$ = \underbar{\bf 50}
\vskip 5pt 
\item{} 90$^{o}$ butterfly valve with offset seat, 20 inch pipe size; $C_v$ = \underbar{\bf 11,600}
\end{itemize}

\vskip 10pt

Note: the $C_v$ values obtained using relative flow coefficients ($C_d$) are {\it approximate only!}  This calculation technique should only be used to {\it estimate} the valve size needed for a particular application.

%(END_ANSWER)





%(BEGIN_NOTES)

The equation students will need to use to solve for $C_v$ in each case, of course, is this:

$$C_v = d^2 C_d$$

\vskip 10pt

Source for $C_d$ factors: \underbar{Chapter 4.17: Valve Sizing}; {\it Instrument Engineer's Handbook, Process Control, Third Edition}; pg. 590.

%INDEX% Final Control Elements, valve: relative flow capacity (defined)
%INDEX% Final Control Elements, valve: relative flow coefficient (defined)
%INDEX% Final Control Elements, valve: sizing

%(END_NOTES)


