
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

A proportional-only controller in automatic mode has the following input and output values:

\begin{itemize}
\item{} PV = 65\%
\item{} SP = 62\%
\item{} Output = 48\%
\end{itemize}

Suddenly, the operator changes the setpoint from a value of 62\% to a value of 55\%.  The controller output immediately goes from 48\% to 31\%.  Calculate the proportional band and the gain for this controller, and show all your work.  Also, determine if this controller is {\it direct} or {\it reverse} acting.

\vfil

\underbar{file i01493}
\eject
%(END_QUESTION)





%(BEGIN_ANSWER)

This is a graded question -- no answers or hints given!

%(END_ANSWER)





%(BEGIN_NOTES)

Just like an electronic amplifier, the {\it gain} of a loop controller is mathematically defined as the ratio of output change ($\Delta m$) to input change ($\Delta e$).  In other words:

$$\hbox{Gain} = {\Delta \hbox{ Output} \over \Delta \hbox{ Input}}$$

$$K_p = {\Delta m \over \Delta e}$$

For this reason, we must approach this problem by examining how far the error {\it changes} from its previous value, and compare that to how far the output signal {\it changes} from its previous value.  In this example, the setpoint changed from 62\% to 55\%, which is a jump of -7\%.  As a consequence of that setpoint change, the output jumped from 48\% to 31\%, which is a jump of -17\%.  Therefore:

$$K_p = {\Delta m \over \Delta e} = {31\% - 48\% \over 55\% - 62\%} = {-17\% \over -7\%} = 2.429$$

Proportional band is defined as the mathematical reciprocal of controller gain, expressed as a percentage.  Therefore:

$$\hbox{PB} = {1 \over K_p} = {1 \over 2.429} = 0.4118 = 41.18\%$$

\vskip 10pt

The fact that the output value decreased when the setpoint value decreased tells us this controller is {\it reverse-acting}.  At first, this may seem backward to us, because didn't the output go in the {\it same} direction as the input?  While this is true, it is imperative to remember that ``direct'' and ``reverse'' for a loop controller is defined as the effect that the {\it process variable} (not the {\it setpoint}) has on the output.  Since we know PV and SP always have opposing effects on the output, a decreasing output for a decreasing SP means the output would increase for a decreasing PV, making it reverse-acting.  The equation for such a controller is shown here:

$$m = K_p (\hbox{SP} - \hbox{PV}) + b$$

%INDEX% Control, proportional: gain versus proportional band
%INDEX& Control, proportional: proportional band versus gain

%(END_NOTES)


