
%(BEGIN_QUESTION)
% Copyright 2006, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

What does the term {\it rangeability} mean, with regard to control valves?  Why is this parameter important to us?  

\vskip 10pt

Which valve opening characteristic exhibits the greatest rangeability: {\it quick opening}, {\it linear}, or {\it equal percentage}?  Explain your answer.

\underbar{file i01423}
%(END_QUESTION)





%(BEGIN_ANSWER)

{\it Rangeability} of a measuring instrument refers to the fraction of full-scale process measurement may be made within the advertised accuracy.  In reference to a control valve, ``rangeability'' refers to the fraction or ratio of full flow that the process fluid flow rate may be reliably controlled to.
 
Rangeability is important because it limits the practical working range of a process flow system.  If a valve has insufficient rangeability, the process flow rate may not be ``turned down'' very much from full flow and still be able to maintain good-quality control.

\vskip 10pt

Equal percentage valve characterization exhibits the best rangeability, because this opening characteristic maintains the finest degree of control toward the lower end of the valve's stroke than any of the others.

Contrast the performance of an equal percentage valve with one that is quick opening, at low position settings: the quick opening valve will be very sensitive to stem/shaft position at the low end (very small movements resulting in large flow changes), while the equal percentage valve will retain good control characteristics toward the bottom end of its range.  This is even more true in process systems with large amounts of characteristic ``distortion'' (resulting from large changes in valve $\Delta$P over the flow range), where all inherent valve characteristics tend to be ``bent'' toward the quick-opening characteristic.
 
%(END_ANSWER)





%(BEGIN_NOTES)



%INDEX% Final Control Elements, valve: rangeability

%(END_NOTES)


