
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Cable Routing'' subsection of the ``Electrical Signal and Control Wiring'' section of the ``Instrument Connections'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i01702}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Electrical conductors must have their weight properly supported, and be protected against damage from external forces.  Several routing techniques exist:

\begin{itemize}
\item{} Conduit
\begin{itemize}

\item{} Plastic (PVC) -- best resistance to corrosion
\end{itemize}
\item{} Liquid-tight flexible conduit
\begin{itemize}

\item{} Nonmetallic -- nothing more than a plastic tube
\end{itemize}
\item{} Cable tray (requires specially-rated cable for exposure)
\item{} Busway -- heavy copper bus conductors inside of sheet metal ducts
\item{} Wire duct -- plastic trays attached inside of enclosures
\item{} Wire loom -- plastic tubes protecting bundles of wires inside enclosures
\end{itemize}

Nylon ``pull string'' used to pull new wires into conduit.  Always leave a length of pull string inside every conduit for future work!

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify an advantage of using metal conduit instead of plastic conduit.
\item{} Identify an advantage of using plastic conduit instead of metal conduit.
\item{} Explain how wire pulling may be made easier by reducing friction between the conductors and the conduit.
\item{} How does metallic liquid-tight flexible conduit differ from non-metallic liquid-tight flexible conduit?
\item{} Identify an advantage of using cable trays instead of conduit.
\item{} Identify an advantage of using conduit instead of cable trays.
\item{} Locate a photograph of industrial cable tray, and describe what the photo shows.
\item{} Locate a photograph of industrial bus duct, and describe what the photo shows.
\item{} Locate a photograph showing wire duct used inside an industrial panel, and describe what the photo shows.
\item{} Locate a photograph showing wire loom used inside an industrial panel, and describe what the photo shows.
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, cable routing

%(END_NOTES)

