
%(BEGIN_QUESTION)
% Copyright 2007, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

List at least five appropriate informational interview questions:

\begin{itemize}
\item{} 
\vskip 30pt
\item{} 
\vskip 30pt
\item{} 
\vskip 30pt
\item{} 
\vskip 30pt
\item{} 
\vskip 30pt
\end{itemize}

\underbar{file i01854}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

\begin{itemize}
\item{} What are the most significant characteristics of this industry?
\item{} What are the growth markets of this industry?
\item{} In what ways is your industry changing?
\item{} How is the economy affecting this industry?
\item{} Does the company have plans for expansion (new buildings, facilities, etc.) in the near future?
\item{} What is the employment outlook like in your specific industry?  How much demand is there for instrument technicians in this company?
\item{} Can you estimate future job openings for instrument technicians?
\item{} How quickly is this industry growing?
\item{} What obstacles do you see getting in the way of the company's profitability or growth?
\item{} Is this industry heavily regulated?
\item{} What are the typical entry-level job titles and functions?
\item{} What is the typical entry-level salary for an instrument technician in this company?
\item{} What are the salary ranges for higher levels?
\item{} Does this company offer a 401K?  If so, is it matching?
\item{} Is there opportunity for advancement into engineering and/or management?
\item{} Is there a salary ceiling?
\item{} What do you find unique about this industry?
\item{} What skills or personal characteristics do you feel contribute most to success in this industry?
\item{} What trends in the field would be most likely to affect someone just entering this career now?
\item{} What kinds of people experience the greatest success in this field?
\item{} What is the most important thing that someone planning to enter this career should know?
\item{} What are the typical working hours for an instrument technician?
\item{} Is shift work involved?  If so, what is the typical shift schedule?
\item{} Do technicians work weekends?
\item{} Are technicians expected to be on call for off-hours problems?
\item{} Do the technicians do most of their work independently or within teams?
\item{} How much travel is involved for technicians at this company?
\item{} What other types of credentials or licenses are required for technicians?
\item{} What do you feel is the best educational preparation for this career?
\item{} Have you heard of Bellingham Technical College, and the Instrumentation program here?
\item{} Does this company offer on-site training to upgrade technicians' skills?
\item{} Does the company offer tuition reimbursement for employees pursuing advanced degrees?
\item{} How does the company recognize outstanding accomplishments of its employees?
\item{} How does the company foster innovation and creativity in its employees?
\item{} What is the average length of time for an employee to stay in the job you hold?
\item{} What incentives or disincentives are there for staying in the same job?
\item{} What practical experiences would help a new graduate qualify for an entry-level technician position (jobshadowing, internships, etc.)?
\item{} Does this company offer jobshadow or internship opportunities to interested students?
\item{} Will this company help technicians maintain professional memberships in organizations such as the ISA (Instrumentation, Systems, and Automation society)?
\item{} Do you have any written job descriptions of positions in this industry/company?
\item{} What is the best way to apply for a technician position at this company? 
\item{} If I wanted to obtain a job here, what would be the best way to learn of job vacancies?
\item{} If you were conducting a job search today, how would you go about it?
\item{} If I wanted to obtain a job here, who would the best person to contact?
\item{} What areas of the company would be most interested in hiring people with my background?
\item{} Would you be willing to answer more questions, by phone or in person, if I need additional advice in the future?
\end{itemize}

%INDEX% Career, job search: informational interview

%(END_NOTES)


