
%(BEGIN_QUESTION)
% Copyright 2007, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Each Ethernet device manufactured in the world today possesses a unique identifying number, known as a {\it MAC address} or {\it hardware address}.  This address is 6 bytes, or octets, long (48 bits).  An example of a valid Ethernet MAC address is shown here:

\vskip 10pt

\centerline{\tt D2-48-1C-30-EA-B5}

\vskip 10pt

Given the number of bits in a MAC address field, how many unique identifier addresses can exist in the world?

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Ethernet MAC addresses are very interesting, in that each one is unique to a single manufactured device.  Explain why this is.
\item{} Describe what might happen in an Ethernet network if two or more devices shared the same MAC address.
\item{} In cyber-crime investigations, MAC addresses can be important.  Explain why, and also describe how the MAC address of a cyber-criminal might be identified.
\item{} On your personal computer, open up a command-line interface (``cmd'' on Microsoft Windows operating systems) and issue the command {\tt arp -a} to see a listing of known IP addresses and their corresponding MAC addresses on your computer's network.
\end{itemize}

\underbar{file i02205}
%(END_QUESTION)





%(BEGIN_ANSWER)

I'll let you figure out the answer to this question, as well as explain how you got it!

%(END_ANSWER)





%(BEGIN_NOTES)

The answer is $2^{48} = 2.81 \times 10^{14}$ addresses, or 281 {\it trillion} addresses.  The first 24 bits of each MAC address is specified by the IEEE Standards Association (IEEE-SA).  A unique 24-bit number is given to each manufacturer by the IEEE-SA, and is called an OUI (Organizationally Unique Identifier).  Each manufacturer then internally assigns the last 24 bits of each MAC address, being careful not to manufacture more than one device with the same address.

\vskip 10pt

This organizational structure yields 16.8 million OUI's for manufacturers, and 16.8 million unique addresses within each OUI.

%INDEX% Networking, Ethernet: MAC address

%(END_NOTES)


