
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

In the United States, electrical enclosures are usually rated for different environmental services by a {\it NEMA Type Designation}.  Research what this type designation means, in general, as well as in particular for the following NEMA types:

\begin{itemize}
\item{} NEMA 1:
\vskip 5pt 
\item{} NEMA 4:
\vskip 5pt 
\item{} NEMA 4X:
\vskip 5pt 
\item{} NEMA 6:
\vskip 5pt 
\item{} NEMA 7:
\vskip 5pt 
\item{} NEMA 8:
\end{itemize}

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify your source of information on NEMA panel designations.
\end{itemize}

\underbar{file i02286}
%(END_QUESTION)





%(BEGIN_ANSWER)

\noindent
{\bf Partial answer:}

\begin{itemize}
\item{} NEMA 4: {\it Intended for indoor or outdoor use, including ``washdown'' with a water hose.}
\vskip 5pt 
\item{} NEMA 6: {\it Same as NEMA 4, but also protects against temporary submersion at a limited depth.}
\end{itemize}

%(END_ANSWER)





%(BEGIN_NOTES)

\begin{itemize}
\item{} NEMA 1: {\it Intended for indoor use, protects against incidental contact and some falling dirt.}
\vskip 5pt 
\item{} NEMA 4: {\it Intended for indoor or outdoor use, including ``washdown'' with a water hose.}
\vskip 5pt 
\item{} NEMA 4X: {\it Same as NEMA 4, but also includes protection against corrosion.  The enclosure may be made of stainless steel or fiberglass instead of regular (mild) steel.}
\vskip 5pt 
\item{} NEMA 6: {\it Same as NEMA 4, but also protects against temporary submersion at a limited depth.}
\vskip 5pt 
\item{} NEMA 7: {\it Intended for indoor hazardous (explosive) atmospheres, Class I, Groups A, B, C, and D as defined in the NEC.}
\vskip 5pt 
\item{} NEMA 8: {\it Intended for indoor or outdoor hazardous (explosive) atmospheres, Class I, Groups A, B, C, and D as defined in the NEC.}
\end{itemize}

\vskip 10pt

An alternative enclosure rating system used outside the United States is the {\it IP} system.  ``IP'' stands for {\it International Protection}, and it is a two-digit numerical code.  The first digit describes the degree of protection against solid objects (people, tools, wires, dust), while the second digit describes the degree of protection against liquids.  The larger the digit (in both cases), the greater the level of protection.


%INDEX% Good practices, wiring: proper NEMA rating for enclosures

%(END_NOTES)


