
%(BEGIN_QUESTION)
% Copyright 2003, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Calculate the full-load line current for a three-phase motor, given a horsepower rating of 150 HP, an efficiency of 93\%, and a line voltage of 480 volts.  Provide one answer for full-load line current assuming perfect power factor (1), and another answer for full-load line current assuming a power factor of 0.90.

\vfil 

\underbar{file i02293}
\eject
%(END_QUESTION)





%(BEGIN_ANSWER)

This is a graded question -- no answers or hints given!

%(END_ANSWER)





%(BEGIN_NOTES)

The direct equivalence between horsepower and watts is 746 watts to 1 horsepower.  Therefore, a 150 HP motor ideally consumes 111,900 watts (perfect efficiency).  With an energy conversion efficiency of 93\%, though, this input power figure will be larger than 111,900 watts to account for energy wasted in heat.  Thus, our actual input power requirement for this 150 HP motor is 120,322.5 watts.

\vskip 10pt

Applying the standard three-phase power formula, we may solve for line current assuming a perfect power factor:

$$P_{total} = \sqrt{3} (I_{line}) (V_{line})$$

$$I_{line} = {P_{total} \over \sqrt{3} (V_{line})}$$

$$I_{line} = {120322.5 \over \sqrt{3} (480)} = 144.73 \hbox{ amps}$$

\vskip 10pt

Now, a power factor of less than 1 means some of the line current will {\it not} be performing useful work, but instead will be involved with the alternating exchange of energy stored and released by the load's inductance.  With a power factor figure of 0.9, only 90\% of the line current will be doing real work.  We may deal with this percentage figure in much the same way as we dealt with efficiency: dividing the ideal line current by this figure will yield the actual line current for the motor:

\vskip 10pt

$I_{line}$ = 160.81 amps (at a power factor of 0.9)

%INDEX% Electronics review: 3-phase voltage/current/power calculation
%INDEX% Electronics review: power factor

%(END_NOTES)


