
%(BEGIN_QUESTION)
% Copyright 2015, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

For each of the following level switch types, determine the applicability to liquid versus solid (granular or powder) process materials, marking the appropriate box with a ``check.''  Be prepared to defend your answers with reasons based on the operating principle of each switch!  Note that you may need to qualify your answers, such as in the case where a particular type of level switch can only detect certain types of liquids or certain types of solids:

% No blank lines allowed between lines of an \halign structure!
% I use comments (%) instead, so that TeX doesn't choke.

$$\vbox{\offinterlineskip
\halign{\strut
\vrule \quad\hfil # \ \hfil & 
\vrule \quad\hfil # \ \hfil & 
\vrule \quad\hfil # \ \hfil \vrule \cr
\noalign{\hrule}
%
% First row
{\bf Switch type} & {\bf Solid material} & {\bf Liquid material} \cr
%
\noalign{\hrule}
%
% Another row
Float &  & \cr
%
\noalign{\hrule}
%
% Another row
Rotating paddle &  & \cr
%
\noalign{\hrule}
%
% Another row
Tuning fork &  & \cr
%
\noalign{\hrule}
%
% Another row
Ultrasonic &  & \cr
%
\noalign{\hrule}
%
% Another row
Capacitive &  & \cr
%
\noalign{\hrule}
%
% Another row
Radar &  & \cr
%
\noalign{\hrule}
%
% Another row
Conductivity &  & \cr
%
\noalign{\hrule}
%
% Another row
Optical (refraction) &  &  \cr
%
\noalign{\hrule}
} % End of \halign 
}$$ % End of \vbox


\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify ways each of these level switch types may be ``fooled'' into either falsely detecting the presence of material, or falsely indicating the absence of material.
\end{itemize}

\underbar{file i02332}
%(END_QUESTION)





%(BEGIN_ANSWER)

You may not find direct answers given for each of these switch types in reference manuals or textbooks -- that's to be expected.  The point of this exercise is to {\it reason} through the operating principle of each to determine whether or not it might be able to detect liquids or solids.

%(END_ANSWER)





%(BEGIN_NOTES)

% No blank lines allowed between lines of an \halign structure!
% I use comments (%) instead, so that TeX doesn't choke.

$$\vbox{\offinterlineskip
\halign{\strut
\vrule \quad\hfil # \ \hfil & 
\vrule \quad\hfil # \ \hfil & 
\vrule \quad\hfil # \ \hfil \vrule \cr
\noalign{\hrule}
%
% First row
{\bf Switch type} & {\bf Solid material} & {\bf Liquid material} \cr
%
\noalign{\hrule}
%
% Another row
Float &  & $\surd$ \cr
%
\noalign{\hrule}
%
% Another row
Rotating paddle & $\surd$ & \cr
%
\noalign{\hrule}
%
% Another row
Tuning fork & $\surd$ & $\surd$\cr
%
\noalign{\hrule}
%
% Another row
Ultrasonic & $\surd$ & $\surd$\cr
%
\noalign{\hrule}
%
% Another row
Capacitive & $\surd$ & $\surd$ \cr
%
\noalign{\hrule}
%
% Another row
Radar & $\surd$ & $\surd$\cr
%
\noalign{\hrule}
%
% Another row
Conductivity & (only if conductive) & (only if conductive) \cr
%
\noalign{\hrule}
%
% Another row
Optical (refraction) &  & $\surd$ \cr
%
\noalign{\hrule}
} % End of \halign 
}$$ % End of \vbox

A float would become buried by solid material filling a vessel from the top (unless the float was part of an automatically-retracting system where a winch reeled it up and down to periodically probe for level).

\vskip 10pt

Paddle wheel switches need a firm substance to stop the rotation, which is why a liquid will not work (unless the liquid is {\it really} viscous).

\vskip 10pt

The vibration of a tuning fork will be damped by any mass, solid or liquid.

\vskip 10pt

Ultrasonic wave propagation will be affected by any mass, solid or liquid.  One must be careful, however, if the process material has a tendency to foul (stick to) the probe.

\vskip 10pt

Capacitive probes can detect any substance (solid or liquid) if the permittivity meets a certain minimum threshold.

\vskip 10pt

Radar instruments are able to measure any substance (solid or liquid) if the change in permittivity from air to the process material is substantial enough.  For non-contact radar instruments, the angle of repose in a solid-containing bin may cause scattering, though.

\vskip 10pt

Conductivity switches will work on any conductive material, solid or liquid.

\vskip 10pt

Optical refraction switches only work with clear liquids, not solids.







%INDEX% Switch, level: process material types

%(END_NOTES)

