
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Explain why every H1 FOUNDATION Fieldbus network segment should have at least a primary {\it and} a ``backup'' LAS device.  On a related note, is {\it every} type of FOUNDATION Fieldbus device capable of being configured as an LAS?

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} If a device is configured to be a backup LAS, what communication events does it ordinarily participate in that non-backup devices do not participate in?
\end{itemize}

\underbar{file i02436}
%(END_QUESTION)





%(BEGIN_ANSWER)

Most Fieldbus segments have two LAS-enabled devices: a primary and a backup.  However, the Fieldbus standard allows for up to 8 LAS-enabled devices (one primary and 7 backups!).  Some host systems, though, do not allow the enabling of all 8 LAS-capable devices.  Emerson's DeltaV DCS, for example, only allows for two: the primary and a single backup LAS.

\vskip 10pt

If there is only one LAS and it fails, the Fieldbus segment will stop working.  This is why it is important to have redundant Link Master devices on each H1 segment.  A ``Link Master'' device is nothing more than a basic Fieldbus device with LAS capability.  FOUNDATION Fieldbus H1 supports multiple Link Masters on a network, the one with the lowest-number address automatically becoming the Link Active Scheduler.

\vskip 10pt

The Fieldbus H1 interface card on the host system (usually a DCS) is usually configured as the LAS device in a Fieldbus system, with one of the field instruments serving as the backup LAS.  Note that some Fieldbus instruments do {\it not} have backup LAS capability, so you must be sure to choose your field devices carefully to ensure at least one of them has this capability!

%(END_ANSWER)





%(BEGIN_NOTES)


%INDEX% Fieldbus, FOUNDATION (H1): Link Active Scheduler (LAS)
%INDEX% Fieldbus, FOUNDATION (H1): Link Master devices

%(END_NOTES)


