
%(BEGIN_QUESTION)
% Copyright 2007, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

In addition to supporting a wide array of general-purpose function blocks for control programming, each FOUNDATION Fieldbus device also contains other blocks specific to that device: one {\it Resource Block} and one {\it Transducer Block}.  What purposes do these other blocks serve, and how do they differ from general function blocks such as AO, AI, PID, etc.?

\underbar{file i02444}
%(END_QUESTION)





%(BEGIN_ANSWER)

The Resource Block contains all the specific parameters and characteristics for a field device.  The algorithm implemented by the Resource Block is unique to the device and may vary from manufacturer to manufacturer.

\vskip 10pt

The Transducer Block handles direct I/O interfacing with sensors or output mechanisms on the device.  Examples would include the differential capacitance cell on a Rosemount 3051 differential pressure transmitter; or the I/P, feedback potentiometer, and pressure sensors on a Fisher DVC6000 valve positioner.

\vskip 10pt

At first it may seem as though the Transducer block would be redundant to the AI or AO blocks.  However, it is not.  The AI and AO blocks are {\it generic} function blocks designed to homogeneously connect with other general-purpose function blocks.  The Transducer Block is specific to the particular device.  AI and AO function blocks interface with Transducer blocks to provide a generic connection point to other function blocks.

Transducer blocks are not just essential for basic device operation, but they are also necessary for AO function blocks to generate ``back calculation'' output signals.  For instance, the pressure sensor and stem position feedback potentiometer inputs on a Fisher DVC6000 valve positioner are used to generate the back-calculation signal value coming from any AO function block configured to it.  These inputs are how the positioner ``knows'' the valve is responding.

%(END_ANSWER)





%(BEGIN_NOTES)


%INDEX% Fieldbus, function block: Resource block (in each device)
%INDEX% Fieldbus, function block: Transducer block (in each device)

%(END_NOTES)


