
%(BEGIN_QUESTION)
% Copyright 2007, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Consider a FOUNDATION Fieldbus H1 segment running over type A cable (18 AWG, individually shielded, twisted-pair cable with $Z_0$ = 100 $\Omega$).  Identify five different {\it physical} problems (i.e. not related to programming or software) resulting from poor segment design and/or installation which would cause communication errors during operation.  Note: neither {\it damage} nor {\it failure} counts as an installation-related problem!  Acceptable answers must account for communication errors without assuming any cable or component in the Fieldbus segment is actually broken.

\vskip 50pt 

\begin{itemize}
\item{$1.$}
\vskip 50pt 
\item{$2.$} 
\vskip 50pt 
\item{$3.$} 
\vskip 50pt 
\item{$4.$} 
\vskip 50pt 
\item{$5.$} 
\end{itemize}

\underbar{file i02532}
%(END_QUESTION)





%(BEGIN_ANSWER)

\begin{itemize}
\item{} Main trunk cable too long.
\vskip 5pt 
\item{} Spur cable(s) too long.
\vskip 5pt 
\item{} Missing terminator at host end.
\vskip 5pt 
\item{} Missing terminator at far end of cable.
\vskip 5pt 
\item{} Extra terminator(s) installed anywhere on the segment.
\vskip 5pt 
\item{} Wrong termination resistor impedance
\vskip 5pt 
\item{} Extra terminator(s) installed where they are not needed (i.e. more than 2).
\vskip 5pt 
\item{} Shield conductor grounded at both ends.
\vskip 5pt 
\item{} Shield conductor left ungrounded.
\vskip 5pt 
\item{} Poor connection(s) at terminal block(s).
\vskip 5pt 
\item{} Non-Fieldbus transmitter installed on segment.
\vskip 5pt 
\item{} Inadequate power supply connected to segment
\vskip 5pt 
\item{} Reversed polarity on device (some FF devices are polarity-sensitive!)
\end{itemize}

%(END_ANSWER)





%(BEGIN_NOTES)


{\bf This question is intended for exams only and not worksheets!}

%(END_NOTES)


