
%(BEGIN_QUESTION)
% Copyright 2011, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose a radio antenna receives 7.5 watts of power from the end of a coaxial cable connected to an RF transmitter outputting 8.5 watts.  The cable is rated to have a power loss of $-0.015$ decibels per foot.  Calculate the output power of the transmitter in units of dBW and dBm, and also the length of the cable in inches.

\vskip 10pt

$P_{xmtr}$ = \underbar{\hskip 50pt} dBW

\vskip 10pt

$P_{xmtr}$ = \underbar{\hskip 50pt} dBm

\vskip 10pt

Cable length = \underbar{\hskip 50pt} inches

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} A technique highly recommended for word-problems is to {\it sketch a picture} of the problem and label elements of that picture with the given information.  Do this, and compare your sketch with those of your classmates.  How, specifically, does this aid your problem-solving?
\end{itemize}

\underbar{file i02998}
%(END_QUESTION)





%(BEGIN_ANSWER)

\noindent
{\bf Partial answer:}

\vskip 10pt

$P_{xmtr}$ = \underbar{\bf 39.294} dBm

%(END_ANSWER)





%(BEGIN_NOTES)

The power attenuation (as a ratio) is 7.5 watts divided by 8.5 watts, or 0.8824.  Expressed as a decibel figure, it is $-0.5436$ dB.  Given the cable's loss rating of $-0.015$ dB per foot, the cable must be 36.24 feet long, or 434.86 inches in length.

\vskip 10pt

Expressing transmitter power in dBW and dBm is a simple matter of logarithmic ratios with reference powers of 1 watt and 1 milliwatt, respectively:

$$10 \log \left(8.5 \hbox{ W} \over 1 \hbox{ W}\right) = 9.294 \hbox{ dBW}$$

$$10 \log \left(8.5 \hbox{ W} \over 1 \hbox{ mW}\right) = 39.294 \hbox{ dBW}$$

Note how the reference powers of dBW and dBm differ by three orders of magnitude ($10^3$), and so the dB power values differ by 3 bels (30 decibels).

%INDEX% Electronics review: decibel power calculations

%(END_NOTES)


