
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Skim the ``Digital PID Controllers'' section of the ``Closed-Loop Control'' chapter in your {\it Lessons In Industrial Instrumentation} textbook to specifically answer these questions:

\vskip 10pt

Explain what ``function blocks'' are in a digital controller.

\vskip 10pt

Where do we typically find {\it DDC} control systems used?

\vskip 10pt

What does the term ``distributed'' refer to in a {\it DCS}?

\vskip 10pt

Describe the typical layout of a {\it SCADA} system.

\vskip 10pt

How do {\it FOUNDATION Fieldbus} instruments differ from ``normal'' microprocessor-based field instruments, and how does this unique capability take the concept of ``distributed control'' to the ultimate level?

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Discuss ideas for ``skimming'' a text to identify key points so you do not have to read the whole thing.
\end{itemize}


\underbar{file i02922}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)



\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain, to the best of your ability, what the function blocks do in the example programs for the Siemens 353 controller.
\item{} Explain what ``fan-out'' refers to in a function block language (or in digital logic gate circuitry, for that matter).
\item{} Describe what BACnet and LonWorks is, and how it is similar to Modbus.
\item{} What is an ``RTU'' in a SCADA system, and how does it differ from an MTU?
\item{} How do SCADA systems differ from telemetry systems?
\item{} How does a DCS differ from a typical SCADA system?
\item{} Describe what FOUNDATION Fieldbus is and what it does.
\end{itemize}










\vfil \eject

\noindent
{\bf Prep Quiz:}

FOUNDATION Fieldbus differs from most other control systems in that:

\begin{itemize}
\item{} The field instruments have built-in microprocessors
\vskip 5pt 
\item{} Radio communication eliminates the need for connecting wires
\vskip 5pt 
\item{} Redundancy ensures continued operation in the event of faults
\vskip 5pt 
\item{} Data is communicated digitally (not analog) between instruments
\vskip 5pt 
\item{} The instruments are able to self-diagnose a range of faults
\vskip 5pt 
\item{} Control ``decisions'' are made in the field devices themselves
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, closed-loop control (digital PID controllers)

%(END_NOTES)


