
%(BEGIN_QUESTION)
% Copyright 2007, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Two very important concepts in process control are {\it energy balance} and {\it mass balance}.  Explain what both of these concepts mean, and then identify how {\it Bernoulli's equation} and the {\it continuity equation} are respective examples of each principle.

\vskip 30pt

\noindent
{\bf Bernoulli's equation:}

$$z_1 \rho g + {v_1^2 \rho \over 2} + P_1 = z_2 \rho g + {v_2^2 \rho \over 2} + P_2$$

\vskip 30pt

\noindent
{\bf Continuity equation:}

$$\rho_1 A_1 v_1 = \rho_2 A_2 v_2$$

\underbar{file i02990}
%(END_QUESTION)





%(BEGIN_ANSWER)

``Energy balance'' means that any process in a steady-state condition must input and output the same amount of energy over time: that every single joule of energy must be accounted for.  This is really the same thing as the Law of Energy Conservation in physics, which states that energy cannot be created or destroyed.

\vskip 10pt

``Mass balance'' means that any process in a steady-state condition must input and output the same amount of mass over time: that every single gram of mass must be accounted for.  This is really the same thing as the Law of Mass Conservation in physics, which states that mass cannot be created or destroyed.

%(END_ANSWER)





%(BEGIN_NOTES)

Bernoulli's equation states that the sum of all energies in a fluid stream must remain constant (assuming no loss).  However, we may even account for energy loss in Bernoulli's equation with the addition of another term:

$$z_1 \rho g + {v_1^2 \rho \over 2} + P_1 = z_2 \rho g + {v_2^2 \rho \over 2} + P_2 + E_{loss}$$

\vskip 10pt

The continuity equation states that mass flow at any point in a single fluid stream must remain constant.  Unless the stream branches in different directions, or mass accumulates in a vessel (or is being released from a vessel), mass flow in must equal mass flow out.

%INDEX% Physics, conservation laws: energy balance
%INDEX% Physics, conservation laws: mass balance
%INDEX% Physics, dynamic fluids: Bernoulli's equation
%INDEX% Physics, dynamic fluids: continuity equation

%(END_NOTES)


