
%(BEGIN_QUESTION)
% Copyright 2014, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose you need to take a current measurement in a CT circuit, inserting an ammeter into the CT's secondary circuit to do so.  The CT's signal passes through a standard ``shorting'' test switch before going to the panel instrument, and one pole of each two-pole CT switch has a ``test jack'' to facilitate current measurement.

\vskip 10pt

Explain how a ``test jack'' on a CT test switch works, and then outline a step-by-step procedure for {\it safely} measuring current in the CT secondary circuit:

\begin{itemize}
\item{} 
\vskip 10pt
\item{} 
\vskip 10pt
\item{} 
\vskip 10pt
\item{} 
\end{itemize}

\underbar{file i03047}
%(END_QUESTION)





%(BEGIN_ANSWER)

Remember that your personal safety depends on getting the procedure correct.  \underbar{You} must thoroughly understand the procedure and why it is necessary in order to keep yourself safe under all conditions!
 
\begin{itemize}
\item{} Connect an AC ammeter to a suitable test plug (matching the test jack).
\vskip 10pt
\item{} Use another meter to test continuity through the ammeter, taking measurements from the blades of the test plug in order to check continuity of the test plug, test leads, and ammeter as one continuous circuit.  It is recommended you move and flex the test leads during this procedure to check for any intermittent ``open'' faults.
\vskip 10pt 
\item{} Open the knife switch where the test jack is located.  Leave the shorting test switch in the closed position so that current still passes through the switch and through the sensing instrument.
\vskip 10pt
\item{} Insert the test plug into the test jack and measure current.
\end{itemize}

%(END_ANSWER)





%(BEGIN_NOTES)


%INDEX% Electronics review: current transformer (CT)
%INDEX% Safety, electrical: proper test plug usage on current transformer (CT) circuit

%(END_NOTES)


