
%(BEGIN_QUESTION)
% Copyright 2014, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Explain what {\it AC reactance} is, and how it differs from {\it resistance} despite being measured in the same units (ohms).  Then, calculate reactance values for the following components:

\begin{itemize}
\item{} 45 millihenry inductor at 400 Hz = \underbar{\hskip 50pt} $\Omega$
\vskip 10pt
\item{} 12 microfarad capacitor at 50 Hz = \underbar{\hskip 50pt} $\Omega$
\vskip 10pt
\item{} Brown-Red-Orange resistor at 1000 Hz = \underbar{\hskip 50pt} $\Omega$
\vskip 10pt
\item{} 1.5 henry inductor at 440 Hz = \underbar{\hskip 50pt} $\Omega$
\end{itemize}

After you have done this, comment on the amount of {\it impedance} exhibited by each of these components.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} One way to think of electrical resistance and electrical impedance in analogous terms is to liken then to mechanical {\it friction} and mechanical {\it inertia}, respectively.  Explain this association in terms of energy transfer.
\item{} Will a series-connected inductor serve as a {\it low-pass} filter or a {\it high-pass} filter?  Explain your answer.
\item{} Explain how to build a {\it low-pass} filter circuit using a capacitor and a resistor.
\end{itemize}

\underbar{file i03066}
%(END_QUESTION)





%(BEGIN_ANSWER)

Resistance is to reactance in an AC electrical circuit as friction is to inertia in a mechanical system.  The former resists motion, while the latter resists {\it changes} in motion.

$$X_L = 2 \pi f L$$

$$X_C = {1 \over 2 \pi f C}$$

$$X_R = 0 \hskip 10pt \hbox{For a ``pure'' resistor}$$

Impedance is the combination of resistance and reactance for any component or network.

%(END_ANSWER)





%(BEGIN_NOTES)

\begin{itemize}
\item{} 45 millihenry inductor at 400 Hz = \underbar{\bf 113.10} $\Omega$
\vskip 10pt
\item{} 12 microfarad capacitor at 50 Hz = \underbar{\bf 265.26} $\Omega$
\vskip 10pt
\item{} Brown-Red-Orange resistor at 1000 Hz = \underbar{\bf 0} $\Omega$
\vskip 10pt
\item{} 1.5 henry inductor at 440 Hz = \underbar{\bf 4.147 k} $\Omega$
\end{itemize}

Impedance will be identical for all components except for the resistor.  A Brown-Red-Orange resistor has a resistance of 12000 ohms, and therefore an impedance of 12000 ohms as well.

%INDEX% Electronics review: AC reactance and impedance

%(END_NOTES)


