
%(BEGIN_QUESTION)
% Copyright 2011, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose a radio transmitting antenna radiates 45 watts of RF energy in a particular direction, and a receiving antenna intercepts 1.4 microwatts of RF energy from it.  Calculate the {\it path loss} between these two antennas, and also calculate the approximate distance between them assuming the frequency of the RF signal is 700 kHz.

\vskip 10pt

$L_{path}$ = \underbar{\hskip 50pt} dB

\vskip 10pt

$D$ = \underbar{\hskip 50pt} meters

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain what creates path loss.  Is this something that can be overcome or at least mitigated by applying the correct technology?  Why or why not?
\item{} Explain what {\it fade loss} is, what creates it, and how it differs from path loss.
\end{itemize}

\underbar{file i03140}
%(END_QUESTION)





%(BEGIN_ANSWER)

Dividing RF power received by RF power transmitted gives us the total path loss as a ratio:

$$L_{path} = {1.4 \> \mu\hbox{W} \over 45 \hbox{ W}} = 3.111 \times 10^{-8}$$

Taking the common logarithm of this figure and multiplying by ten yields the same path loss as a decibel figure:

$$L_{path} = 10 \log (3.111 \times 10^{-8}) = -75.07 \hbox{ dB}$$

\vskip 10pt

Now, we know path loss in decibels is related to path distance through empty space by the following formula:

$$L_{path} = -20 \log \left(4 \pi D \over \lambda \right)$$

All we need to do now is solve for the wavelength of the RF signal ($\lambda$) and then manipulate the path loss formula to solve for distance ($D$):

$$v = \lambda f$$

$$\lambda = {v \over f} = {2.9979 \times 10^{8} \hbox{ m/s} \over 700000 \hbox{ Hz}} = 428.27 \hbox{ m}$$

\vskip 10pt

$$L_{path} = -20 \log \left(4 \pi D \over \lambda \right)$$

$${L_{path} \over -20} = \log \left(4 \pi D \over \lambda \right)$$

$$10^{L_{path} \over -20} = {4 \pi D \over \lambda}$$

$$\lambda 10^{L_{path} \over -20} = 4 \pi D$$

$$D = {\lambda 10^{L_{path} \over -20} \over 4 \pi}$$

$$D = {(428.27) \left(10^{-75.07 \over -20} \right) \over 4 \pi} = 193,219.7 \hbox{ m} = 193.22 \hbox{ km}$$

\vskip 10pt

Note: the wavelength of this RF signal (428.27 meters) is typical of AM radio transmissions, which is why the transmitting antennas for AM radio stations must be so tall!  Antenna size is a direct function of signal wavelength, in order that the signal will resonate with the physical length of an antenna the same way that acoustic (sound) waves resonate inside the length of a wind instrument such as a saxophone or trumpet.

%(END_ANSWER)





%(BEGIN_NOTES)

%INDEX% Electronics review: decibel power calculations

%(END_NOTES)


