
%(BEGIN_QUESTION)
% Copyright 2012, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

An engineer designing an electric crane for lifting heavy loads decides to use VFD controls on all the winch motors, for precise control of lifting and dropping speeds, as well as the ability to put electrical power back into the grid using {\it regenerative} braking.  

The engineer also decides to configure the winch control VFDs as safety brakes, to slow down the cables' release in the event that speed switches sensing rotation on any of the cable winch drums detect a rapidly falling load.  His idea is to have the crane automatically detect if the load is dropped too quickly, and for the VFDs to brake this descent automatically without human operator intervention.

\vskip 50pt

First, describe a scenario where this electric crane will output power to the grid.  In other words, under what operation condition(s) will the VFDs regenerate as intended?

\vskip 100pt

Second, identify a potential problem with engineer's decision to use the VFDs as safety brakes for arresting a falling load.

\underbar{file i03406}
%(END_QUESTION)





%(BEGIN_ANSWER)

Each VFD will regenerate when it is acting as a brake: that is to say, when the load is being lowered, not raised.

\vskip 10pt

Using the VFDs as automatic safety brakes will work if and only if the crane does not lose line power.  If the crane drops a load due to a lack of incoming AC power, the VFDs will shut off and not be able to provide any braking action, regenerative or not.  Several other critiques of using a VFD as a safety brake include:
 
\begin{itemize}
\item{} It cannot fully stop a load, only slow it down
\item{} The VFD could lose its digital parameters and therefore compromise safety
\item{} This will add more heat to a (possibly) already overheated motor
\item{} If the motor fails, the brake fails
\item{} May require a ridiculously oversized VFD to provide adequate braking horsepower
\item{} Emergency braking may tax power grid in the location of the crane
\end{itemize}

%(END_ANSWER)





%(BEGIN_NOTES)

{\bf This question is intended for exams only and not worksheets!}.

%(END_NOTES)


