
%(BEGIN_QUESTION)
% Copyright 2011, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read the whitepaper published by Rosemount on the topic of top-mounting DP flowmeters on steam lines (``Top Mount Installation for DP Flowmeters in Steam Service'', document 00870-0200-4809, copyright August 2009) and answer the following questions:

\vskip 30pt

Why has the traditional recommendation for DP flow transmitter on steam lines been to locate the transmitter {\it below} the line?

\vskip 50pt

What kind(s) of problem(s) are typically experienced with below-pipe mounting of DP flow transmitters in steam line applications?

\vskip 50pt

Can DP flowmeters {\it always} be top-mounted?  If not, what limitations dictate whether or not to top-mount?

\vskip 50pt

Why shouldn't Annubar-style flow elements be mounted {\it vertically} in a steam pipe, but rather should be canted at least 15 degrees from vertical?

\underbar{file i03488}
%(END_QUESTION)





%(BEGIN_ANSWER)

Why has the traditional recommendation for DP flow transmitter on steam lines been to locate the transmitter {\it below} the line?  {\it Below-line mounting in steam service helps protect the transmitter against damage from high steam temperatures.}

\vskip 10pt

What kind(s) of problem(s) are typically experienced with below-pipe mounting of DP flow transmitters in steam line applications? {\it Measurement errors at low DP values due to uneven water columns in ``wet leg'' impulse lines.  The water in the wet impulse legs can also freeze in cold weather.}

\vskip 10pt

Can DP flowmeters {\it always} be top-mounted?  If not, what limitations dictate whether or not to top-mount?  {\it Top-mounting is applicable only for certain limited temperature ranges.  Otherwise, the pipe is simply too hot and the transmitter will be ``cooked'' to death.}

\vskip 10pt

Why shouldn't Annubar-style flow elements be mounted {\it vertically} in a steam pipe, but rather should be canted at least 15 degrees from vertical?  {\it To avoid measurement errors due to water running alongside the bottom of the steam line, impacting the lowest port on the Annubar element.}

%(END_ANSWER)





%(BEGIN_NOTES)



%INDEX% Reading assignment: Emerson/Rosemount whitepaper (Top Mount Installation for DP Flowmeters in Steam Service)

%(END_NOTES)


