
%(BEGIN_QUESTION)
% Copyright 2013, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose a PLC sends Modbus command signals to a variable-frequency motor drive (VFD) over a serial EIA/TIA-232 network cable.  Both the PLC and the VFD are equipped with standard 9-pin serial ports, with the following connections between serial port pins and cable conductors:

\begin{itemize}
\item{} Pin 2 on VFD (RD) $\rightarrow $ red wire of serial cable  Pin 3 on PLC (TD)
\item{} Pin 3 on VFD (TD) $\rightarrow $ green wire of serial cable  Pin 2 on PLC (RD)
\item{} Pin 5 on VFD (Gnd) $\rightarrow $ black wire of serial cable  Pin 5 on PLC (Gnd)
\item{} Shield conductor of serial cable  chassis ground at PLC
\end{itemize}

You intend to connect a personal computer (also equipped with a 9-pin serial port) to this network in such a way that it is able to ``listen'' to the Modbus replies coming from the VFD.  Your plan is to use {\tt Hyperterminal} to display the Modbus codes in a format you can read on the computer screen.

\vskip 10pt

Specify which pins of the personal computer's serial port need to connect to which wires of the serial cable in order to allow it to listen to the VFD's Modbus responses to the PLC.  Also, identify whether the PLC/VFD Modbus communications need to be in {\it RTU} or {\it ASCII} format.

\underbar{file i03544}
%(END_QUESTION)





%(BEGIN_ANSWER)

3 points for each pin/wire connection, 4 points for correct Modbus format.  Deduct 2 points if the student has connected Pin 3 of the computer (TD) to any wire of the cable (since there is no need for the computer to transmit on this network, and doing so may cause problems).

\vskip 10pt

Pin 2 on the computer (RD) needs to connect to the green wire.

Pin 5 on the computer (Gnd) needs to connect to the black wire.

The Modbus communication needs to be ASCII in order for {\tt Hyperterminal} to display the data.

%(END_ANSWER)





%(BEGIN_NOTES)

{\bf This question is intended for exams only and not worksheets!}.

%(END_NOTES)


