
%(BEGIN_QUESTION)
% Copyright 2011, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose a small city water department happens to upgrade their old leased-line based SCADA system to use 900 MHz industrial radio transceiver units and antennas instead for relaying water storage tank data (water level, pump status, temperature, etc.) at tank locations all over town to the main control center located in the geographic center of town.  After installing the new radio units and antennas, it is noticed that some of the water tank RTU sites are registering weak signal reception while others are within healthy operating limits.

As a solution to the weak-signal problem, one technician suggests replacing the 1/2 wave ``whip'' antenna at the main control center with a Yagi antenna having a much greater gain.  Another technician disagrees, suggesting instead to upgrade each antenna at all the weak-signal locations from ``whip'' style to Yagi.

\vskip 20pt

What do you think?  Do you agree with one of these technicians more than the other?  If so, why?

\vskip 100pt

Also, suggest \underbar{two} solutions {\it other} than antenna replacement, either one of which would independently solve the problem in this SCADA system.  Note that this is {\it not} a mesh-networking system like WirelessHART, but rather a point-to-point radio network where each RTU is independent of the others.

\vskip 100pt

\underbar{file i03644}
%(END_QUESTION)





%(BEGIN_ANSWER)

4 points for choosing the best technician's idea and explaining why, and 3 points for each correct alternative solution:

\vskip 10pt

The first technician's solution (replacing the MTU whip with a Yagi) is flawed, because Yagi antennas are directional and the MTU needs an omnidirectional antenna.  The second technician's idea of replacing individual RTU antennas with Yagis is better.
 
\vskip 10pt

Non-antenna-replacement solutions include more powerful/sensitive radio transceivers, a taller central antenna tower, installing repeaters between weak RTUs and base, or even reverting to leased-lines for some of the weakest RTU locations (keeping radio for the others).

%(END_ANSWER)





%(BEGIN_NOTES)

{\bf This question is intended for exams only and not worksheets!}.

%(END_NOTES)


