
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Use of Line Reactors'' section of the ``Variable-Speed Motor Controls'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i03841}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

DC and AC motor drives alike produce lots of electrical noise due to the PWM switching of their power transistors.  Fourier proved that any cyclic waveform is equivalent to a sum of harmonic frequencies, therefore motor drives produce harmonics on their power lines.  These harmonics may cause transformer core heating, resonance with power-factor correction capacitors, and even conductor overheating in the case of three-phase systems where triplen harmonics (3rd, 6th, 9th, 12th) add in the neutral of Wye-connected systems and circulate in the phase elements of Delta-connected systems.

\vskip 10pt

Series inductors may be used as low-pass filters to keep harmonics contained to the VFD.  Such line reactors may be placed in the drive's input conductors as well as on the drive's output conductors.  Line reactors also help to limit fault currents in the event of short-circuited wiring!








\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain how reactors help mitigate problems caused by harmonic frequencies generated by motor drives.
\item{} Explain why one might choose to install an input line reactor in a VFD system.  What, specifically, is being protected by this reactor?
\item{} Explain why one might choose to install an output line reactor in a VFD system.  What, specifically, is being protected by this reactor?
\item{} Will the presence of a line reactor affect a VFD's ability to {\it brake} the motor using DC injection? 
\item{} Will the presence of a line reactor affect a VFD's ability to {\it dynamically brake} the motor? 
\item{} Will the presence of a line reactor affect a VFD's ability to {\it regeneratively brake} the motor? 
\item{} Will the presence of a line reactor affect a VFD's ability to {\it plug} the motor? 
\item{} Explain what will happen to a VFD-controlled motor if one of the inductors in an input line reactor were to fail open.
\item{} Explain what will happen to a VFD-controlled motor if one of the inductors in an output line reactor were to fail open.
\end{itemize}


%INDEX% Reading assignment: Lessons In Industrial Instrumentation, line reactors in motor speed control circuits

%(END_NOTES)

