
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Calculate the amount of heat energy released by two pounds of superheated steam (at atmospheric pressure) as it cools from 400 $^{o}$F to 125 $^{o}$F, in units of BTU.  Be sure to separate your solution into three steps: the heat lost as the steam cools to the condensing temperature (212 $^{o}$F), the latent heat released through condensation, and the heat lost as the condensed water cools to the final temperature of 125 $^{o}$F.

In which step of this three-step heat loss process is {\it most} of the heat being released?  What does this indicate about the heat-storing capabilities of water, steam, and phase changes between water and steam?

\vskip 10pt

Next, use numbers taken from a {\it steam table} to calculate the same heat release: subtracting the enthalpy of 400 $^{o}$F steam from the enthalpy of 125 $^{o}$F water.  Assume the steam is at atmospheric pressure.  How closely does this calculated value agree with your previous calculation?

\vskip 10pt

Hint: the {\it Socratic Instrumentation} website contains a page where you may download public-domain textbooks, one of which is a set of steam tables published in 1920.  The {\it Fisher Control Valve Handbook} also has a (less comprehensive) set of steam tables in the Appendix section.




\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Be sure to discuss how to properly use a steam table to reference enthalpy values.
\item{} Some superheated steam tables index their entries by the ``total'' temperature of the steam, while others list the amount of ``superheat.''  What exactly does {\it superheat} mean, and how do we calculate it?
\item{} Which result do you think is more accurate: the one calculated in three steps, or the one calculated from steam table enthlapy?
\end{itemize}

\underbar{file i03982}
%(END_QUESTION)





%(BEGIN_ANSWER)

\noindent
{\bf Calculation in three steps:}

\vskip 10pt

\noindent
{\it Heat lost as steam cools from 400 $^{o}$F to 212 $^{o}$F:}

$$Q = mc \Delta T = (2 \hbox{ lb})(0.476 \hbox{ BTU/lb}\cdot^o\hbox{F})(400^o\hbox{F} - 212^o\hbox{F}) = 178.98 \hbox{ BTU}$$

\vskip 10pt

\noindent
{\it Heat lost as steam condenses to water:}

$$Q = mL = (2 \hbox{ lb})(970.3 \hbox{ BTU/lb}) = 1940.6 \hbox{ BTU}$$

\vskip 10pt

\noindent
{\it Heat lost as water cools from 212 $^{o}$F to 125 $^{o}$F:}

$$Q = mc \Delta T = (2 \hbox{ lb})(1 \hbox{ BTU/lb}\cdot^o\hbox{F})(212^o\hbox{F} - 125^o\hbox{F}) = 174 \hbox{ BTU}$$

\vskip 10pt

Total Heat Lost = 178.98 + 1940.6 + 174 = {\bf 2293.58 BTU}

\vskip 30pt

\noindent
{\bf Calculation using enthalpy values (from a steam table):}

\vskip 10pt

\noindent
Enthalpy of 400 $^{o}$F superheated steam at 1 atm = 1239.9 BTU/lb = 2479.8 BTU for two pounds steam

\vskip 10pt

\noindent
Enthalpy of two pounds 125 $^{o}$F water (heat lost cooling down to 32 $^{o}$F):

$$Q = mc \Delta T = (2)(1)(125 - 32) = 186 \hbox{ BTU}$$

\vskip 10pt

Total Heat Lost = 2479.8 $-$ 186 = {\bf 2293.8 BTU}

%(END_ANSWER)





%(BEGIN_NOTES)

The calculation using enthalpy (total heat) values from the steam table is actually more accurate than the three-step calculation because the steam table factors in the (slightly) changing specific and latent heat values of water and steam over a wide temperature range.  Our three-step calculation assumed constant $L$ and $c$ values for all temperatures and pressures.

\vskip 10pt

Of course, the degree of accuracy using a steam table depends on how closely that table lists your precise conditions of temperature and pressure.  The superheated steam table contained in the Fisher {\it Control Valve Handbook} lists the following:

\begin{itemize}
\item{} Absolute pressure = 14.696 PSIA
\item{} Total temperature = 400 $^{o}$F
\item{} Enthalpy ($h_g$) = {\bf 1239.9 BTU/lb}
\end{itemize}

\vskip 10pt

By contrast, {\it Tables and Diagrams of the Thermal Properties of Saturated and Superheated Steam} by Lionel Marks and Harvey Davis (published in 1920) does not contain an entry at these same exact parameters.  Instead, the closest it shows us (on page 25) is the following:

\begin{itemize}
\item{} Absolute pressure = 15 PSIA
\item{} Superheat temperature = 190 $^{o}$F
\item{} Enthalpy ($h$) = {\bf 1239.7 BTU/lb}
\end{itemize}

%INDEX% Physics, heat and temperature: calorimetry problem (with latent heat) 
%INDEX% Physics, heat and temperature: steam table

%(END_NOTES)


