
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Skim the ``Continuous Level Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook to specifically answer these questions:

\vskip 10pt

Explain how nuclear radiation may be used to detect the level of process fluid (or solids) inside a vessel.

\vskip 10pt

Nuclear-type level sensors are often more expensive to install and maintain than other level-measurement technologies.  Given this fact, explain why anyone would choose nuclear over more common technologies for sensing process fluid level.


\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify different strategies for ``skimming'' a text, as opposed to reading that text closely.  Why do you suppose the ability to quickly scan a text is important in this career?
\end{itemize}

\underbar{file i03946}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Nuclear radiation will be blocked (or at least attenuated) by the presence of substance inside a process vessel.  Thus, the intensity of the radiation received will indicate level.

\vskip 10pt

We use these instruments only where no other technology will work.

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Continuous Level Measurement (nuclear)

%(END_NOTES)


