
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Skim the ``Continuous Level Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook to specifically answer these questions:

\vskip 10pt

Explain what a {\it sightglass} is, and how it functions similar to a U-tube manometer.

\vskip 10pt

Explain how a {\it float} may be used to measure liquid level in a vessel, both by hand and automatically.

\vskip 10pt

Identify some potential problems exhibited by sightglasses and floats.


\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify different strategies for ``skimming'' a text, as opposed to reading that text closely.  Why do you suppose the ability to quickly scan a text is important in this career?
\end{itemize}

\underbar{file i03940}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

A {\it sightglass} is a clear tube mounted to the side of a vessel, though which one may direcly view the level of the liquid inside the vessel.  Sightglasses are connected to the vessel both at the top and bottom of the tube, usually through shut-off valves called {\it gauge valves}.  As such, the sightglass tube serves as one-half of a U-tube manometer that always reads zero differential pressure (i.e. the liquid level inside the vessel should always be the same as the liquid level inside the glass tube).

Sightglasses may not properly indicate the level of a liquid-liquid interface unless both ports of the sightglass are fully submerged.  Differing temperatures between glass and vessel may also cause measurement errors.

\vskip 10pt

A {\it float} directly senses liquid level by floating on the surface of the liquid, and transferring this measurement via cable or tape to some instrument head outside the vessel.  Problems include lateral motion from liquid turbulence, fouling of cable or tape.

Some float instruments work like sightglasses: a small float rides on the surface of a liquid column inside of a standpipe, which itself is analogous to a sightglass tube.

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Continuous Level Measurement (sightglasses and floats)

%(END_NOTES)


