
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Valve Manifolds'' subsection of the ``Pressure Sensor Accessories'' section of the ``Continuous Pressure Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.


\underbar{file i03916}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

A valve manifold is a device containing multiple hand valves used to isolate a pressure-sensing instrument from the process.  ``Block'' valves prevent process fluid from making it through to the instrument.  ``Bleed'' valves vent pressure to atmosphere.  ``Equalizing'' valves connect the high and low ports of the instrument together to create a condition of 0 differential pressure.

\vskip 10pt

To isolate a working DP transmitter using a 3-valve manifold:

\item{} Close ``high'' side block valve
\item{} Open equalizing valve
\item{} Close ``low'' side block valve
\item{} Open bleed valve
\end{itemize}

To return this DP transmitter to service, reverse the above procedure.  (``To HEL and back'')

\vskip 10pt

5-valve manifolds typically have two equalizing valves and a vent valve connected between the two equalizing valves.  2-valve manifolds (one block and one bleed) are used in gauge and absolute applications.

\vskip 10pt

It is important to never let live process fluid flow through a valve manifold if that fluid is hot or otherwise dangerous.  In other words, don't open the equalizer valve while both block valves are open.

\vskip 10pt

A good practice is to ``back off'' a hand valve about $1 \over 4$ turn after opening fully, so that it is less likely to seize in that position and feel like a closed valve.




\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain why the valves on a three-valve manifold should be operated in a particular sequence.
\item{} Identify what bad things could happen if the equalizing valve is left open while the block valves are both open.
\item{} Zoom in on the photograph of the eight pressure transmitters, and have students comment on the details they can see.
\item{} What would happen if the equalizer valve of a 3-valve manifold developed a leak and could not be completely shut off?
\item{} What would happen if the ``low'' side block valve of a 3-valve manifold developed a leak and could not be completely shut off?
\item{} What would happen if the ``high'' side block valve of a 3-valve manifold developed a leak and could not be completely shut off?
\item{} What would happen if one of the equalizer valves of a 5-valve manifold developed a leak and could not be completely shut off?
\item{} Explain why it is a good idea to ``back off'' a manifold valve about one-quarter turn after {\it opening} it fully.
\item{} Should you ``back off'' a quarter turn after {\it closing} a manifold valve?  Explain why or why not.
\end{itemize}








\vfil \eject

\noindent
{\bf Prep Quiz:}

Identify the best sequence of 3-valve manifold valve operations when removing a DP transmitter from service.

\begin{itemize}
\item{} Shut the High side block, shut the Low side block, and open the Equalizer
\vskip 5pt 
\item{} Open the Equalizer, shut the Low side block, and shut the High side block
\vskip 5pt 
\item{} Shut the Low side block, shut the High side block, and open the Equalizer
\vskip 5pt 
\item{} Shut the Low side block, open the Equalizer, and shut the High side block
\vskip 5pt 
\item{} Shut the High side block, open the Equalizer, and shut the Low side block 
\vskip 5pt 
\item{} Open the Equalizer, shut the High side block, and shut the Low side block
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Pressure sensor accessories (transmitter manifolds)

%(END_NOTES)


