
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Pressure Switches'' section of the ``Discrete Process Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i03934}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

``Normal'' state is its condition of rest: minimal applied pressure.

\vskip 10pt

Some pressure switches use Bourdon tube as sensing element, tilting a mercury switch to make and break electrical contacts.  Mercury contacts are sealed from atmosphere, but susceptible to vibration and mounting angle.

``Deadband'' or ``differential'' refers to the amount of pressure difference between its trip and its reset pressure values.  This is distinct from the concept of ``differential pressure'' in the cases where the switch has both High and Low pressure ports and actually senses differences of fluid pressure.



\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Define ``normal'' for a pressure switch, as the term is used for NO and NC switch contacts.
\item{} Explain how a mercury tilt-bulb style of switch functions, and why one might be used instead of an open-contact switch.
\item{} For those who have studied classified areas, identify specific applications where a mercury tilt-bulb style of switch might be useful.
\item{} Identify a practical application where a mercury tilt-bulb style of pressure switch would {\it not} be suitable.
\item{} Explain what the ``deadband'' or ``differential'' pressure value means for a pressure switch.
\item{} Describe a scenario where an incorrect ``deadband'' setting for a pressure switch might cause a problem in the instrument system.
\item{} Identify a practical application for a differential pressure switch (the type with two pressure ports to connect to the process).
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Pressure switches

%(END_NOTES)


