
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Heat-Traced Impulse Lines'' subsection of the ``Pressure Sensor Accessories'' section of the ``Continuous Pressure Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.


\underbar{file i03921}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Steam lines and electrical ``heat tape'' may be used to heat impulse lines and instrument flanges to prevent freezing of fluid during cold weather.

\vskip 10pt

{\it Steam traps} allow more steam to enter a steam-tracing line based on local temperature.  They release condensed water according to how quickly that water accumulates in the trap.  The colder the ambient temperature, the faster the rate of condensation, and the more frequently the trap will ``blow'' to drain that water and allow new steam to enter the tracing line.

\vskip 10pt

Electrical heat tape may be self-regulating for temperature, or controlled by an external thermostat.





\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Which portion of a differential pressure transmitter (capsule versus electronics) should be heat-traced, if any?
\item{} Identify ways in which a steam trap might fail, and the potentially adverse effects these faults could have on the pressure instrument.
\item{} Suppose a length of electrical {\it heat tape} suffered an ``open'' fault half-way down its length.  What would happen as a result of this electrical fault?
\end{itemize}


%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Pressure sensor accessories (heat tracing and heated enclosures)

%(END_NOTES)


