
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Extension Wire'' subsection of the ``Thermocouples'' section of the ``Continuous Temperature Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.


\underbar{file i03993}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

If copper wire is used to connect a thermocouple to a thermocouple-reading instrument, the reference junction will be located near the process thermocouple, where the ambient temperature is unknown to the receiving instrument.  Therefore, the receiving instrument's reference junction compensation will be uncertain.

\vskip 10pt

In order to keep the reference junction at the measuring instrument, we must use either thermocouple cable or special {\it extension-grade} cable to connect the thermocouple to the instrument.  Extension grade cable uses inexpensive metal alloys that mimic the true thermocouple metals within the narrow range of temperatures expected at the junction points between the thermocouple and extension wire, and also between the extension wire and the measuring instrument.

\vskip 10pt

The outer sheath of extension wire is colored according to thermocouple type, unlike thermocouple-grade wire which is always plain brown in sheath color.






\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Why is it a bad idea to use copper extension wire between a field-mounted thermocouple and a panel-mounted transmitter?
\item{} Why might it be a bad idea to use thermocouple-grade wire between a field-mounted thermocouple and a panel-mounted transmitter?
\item{} How does extension-grade wire differ from copper or thermocouple-grade wire?
\item{} How do we visually distinguish extension-grade wire from thermocouple-grade wire?
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Continuous Temperature Measurement (thermocouples)

%(END_NOTES)


