%(BEGIN_QUESTION)
% Copyright 2015, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

An ecological survey team installs a Cippoletti weir in a small stream to measure water flow through it.  Calculate the amount of water flow (in units of ft$^{3}$/sec) represented by a crest height (``head'') of 5 inches.  Assume the weir has a crest width of 4 feet and that the crest height is being measured by a level sensor located 3 feet upstream of the weir.  

\vskip 10pt

Also, convert this flow value into units of gallons per minute.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Why do you think a weir would be a good candidate technology for measuring the flow rate of water down a small stream?
\item{} Do you see any ways that a Cippoletti weir could experience problems measuring water flow down a natural stream?  If so, can you think of a better flowmeter technology for this application?
\end{itemize}
  
\underbar{file i04084}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

$$Q = 3.367 L H^{1.5} \hbox{\hskip 20pt Cippoletti weir}$$

\noindent
Where,

$Q$ = Volumetric flow rate (cubic feet per second -- CFS)

$L$ = Width of crest (feet)

$H$ = Head (feet)

\vskip 10pt

$H$ = 5 inches = $5 \over 12$ feet = 0.4167 feet

$$Q = (3.367) (4) (0.4167)^{1.5}$$

$Q$ = 3.622 CFS 

\vskip 10pt

Now, converting 3.622 cubic feet per second into gallons per minute:

$$\left(3.622 \hbox{ ft}^3 \over \hbox{sec} \right) \left(60 \hbox{ sec} \over 1 \hbox{ min}\right) \left(12^3 \hbox{ in}^3 \over 1^3 \hbox{ ft}^3\right) \left(1 \hbox{ gal} \over 231 \hbox{ in}^3\right)$$

$Q$ = 1625.8 GPM

\vskip 10pt

The location of the level-sensing element (3 feet upstream of the weir) is extraneous information, included for the purpose of challenging students to identify whether or not information is relevant to solving a particular problem.

%INDEX% Measurement, flow: weir

%(END_NOTES)


