
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit


Skummles "Continuous Fluid flow Measurment" kapittelet i afgv.pdf for spesifikt å besvare disse spørsmålene:
\vskip 10pt

Forklar hvordan et turbinflowmeter måler hastigheten til et fluid når strømmer igjennom et rør. 

\vskip 10pt


Forklar hva {\it von K\'arm\'an} effekten er, og hvordan den kan utnyttet til å måle hastigheten til  en gass eller en væske. 

\vskip 10pt

Forklar hvordan en elektromagnetisk strømningmåler måler hastigheten til et fluid. 

\vskip 10pt

Forklar hverdan en ultralyd strømningsmåler måler hastigheten til et fluid. 
\vskip 10pt

Explain how an {\it optical} flowmeter measures fluid velocity.
Forklar hvordan et optisk flowmeter måler hastigheten til et fluid. 

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify different strategies for ``skimming'' a text, as opposed to reading that text closely.  Why do you suppose the ability to quickly scan a text is important in this career?
\end{itemize}

\underbar{file i04022}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Turbine flowmeters use a free-spinning ``windmill'' placed in the fluid flow to measure that fluid's velocity.

\vskip 10pt

The von K\'arm\'an effect is the alternating series of vortices formed when a fluid moves past a blunt obstacle.  The frequency of this vortex shedding is the basis of a vortex flowmeter: the higher the fluid's velocity, the faster the oscillation ensues.

\vskip 10pt

Magnetic flowmters exploit the electrical conductivity of the process liquid to generate a small voltage based on fluid velocity.  The fluid moves through a magnetic field, generating an EMF perpendicular to both the field and the direction of flow.

\vskip 10pt

Ultrasonic flowmeters measure fluid velocity by passing high-frequency sound waves through the fluid path.  ``Doppler'' flowmeters bounce sound waves off of bubbles or solids in the liquid stream, inferring velocity from the pitch of the echo.  ``Transit time'' flowmeters send a sound wave upstream and downstream, comparing the times of flight for each pulse.  The difference in time indicates fluid velocity.

\vskip 10pt

Optical flowmeters measure fluid velocity by passing light through the moving fluid and monitoring the effects that fluid has on the light's passage.  In an ``L2F'' (laser-two-focus) optical flowmeter, the sequential interruption of a two light beams by any moving particles in the fluid is used to time the motion of those particles and thereby infer fluid velocity.  In a ``scintillation'' optical flowmeter, turbulence within the clear fluid warps the light beam (the same effect that makes stars appear to ``twinkle'' when viewed from the surface of the Earth), and the sequential detection of common scintillation patterns by a pair of optical sensors is used to time the motion of the fluid and thereby infer fluid velocity.




%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Continuous Fluid Flow Measurement (velocity)

%(END_NOTES)


