
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the introduction of the ``True Mass Flowmeters'' section of the ``Continuous Fluid Flow Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\vskip 10pt

Convert a volumetric flow rate of water equal to 500 CFM (cubic feet per minute) into units of pounds (mass) per second.

\underbar{file i04072}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Many flowmeters sense fluid {\it velocity} (e.g. turbine, ultrasonic, vortex) and as such are volumetric instruments.  However, in applications where we need to measure the molecular quantity of fluid moving through, a flowmeter sensing mass would be more appropriate.

\vskip 10pt

An application of mass flow measurement is chemical feed flow, where we need to know how many molecules of a substance are flowing into a reactor.  It does not matter how densely these molecules are packaged, and so volumetric measurements would be useless.

Another application of mass flow measurement is in steam boilers, to relate water flow in to steam flow out.  What really matters is an accounting of mass, since every pound of water boiled will make a pound of steam.  Volumetric flow measurement on a boiler would be meaningless.

Custody transfer is another application for mass flow measurement, where we are fundamentally concerned with the purchase and sale of molecules.  Since mass is an intrinsic property of matter, we know in advance how much mass each molecule (of a certain type) will have, and so gross measurements of mass tell us how many molecules have passed through.

\vskip 10pt

Mass flow ($W$) and volumetric flow ($Q$) are related to each other by the following formula:

$$W = \rho Q$$

\vskip 10pt

Volumetric flowmeter technologies may be compensated by pressure and temperature measurements to yield a computed mass flow measurement.  This is the principle of AGA3 (orifice plate), AGA7 (turbine), and AGA9 (ultrasonic) gas flow measurement standards for custody transfer.  In each of these cases, the signal output by the flow element must be ``compensated'' to yield a calculated value for mass flow rate, since the flow element itself is not inherently a mass-sensing device.

\vskip 10pt

True mass flowmeters, by contrast, naturally respond to the flow rate in mass units.  Coriolis flowmeters sense the inertia of a flowing fluid, and thermal flowmeters sense the heat transferred by a flowing fluid.

$$ \left(500 \hbox{ ft}^3 \over \hbox{min} \right)  \left(62.4 \hbox{ lb} \over 1 \hbox{ ft}^3\right) = 31200 \hbox{ lb/min} = 520 \hbox{ lb/s}$$










\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Compare and contrast AGA3, AGA7, and AGA9 flow measurement standards.
\item{} Provide at least one industrial flow-measurement example where {\it mass flow} is required rather than {\it volumetric flow}.
\item{} Explain why {\it mass flow} measurement is preferred over {\it volumetric flow} measurement for feedwater and steam flow rates in an industrial boiler.
\item{} Explain why {\it mass flow} measurement is preferred over {\it volumetric flow} measurement for feedstock flow rates in chemical reaction processes.
\item{} Catalyzed paints, such as the types used by professional painters for painting automotive bodies, come in two parts which must be precisely mixed just prior to application.  When mixing these chemical components, the best way to do so is to measure out each of them by weight rather than by volume.  Explain why this is.
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Continuous Fluid Flow Measurement (mass flow measurement)

%(END_NOTES)


