
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Other Differential Producers'' subsection of the ``Pressure-Based Flowmeters'' section of the ``Continuous Fluid Flow Measurement'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i04041}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

{\it Pitot tubes} generate a pressure based on fluid {\it stagnation} at the entrance of the upstream-facing tube.  Traditional Pitot tubes are sensitive to flow at only one point in the stream, so {\it averaging} Pitot tube assemblies are made with rows of holes to average out the effects of the flow profile across the pipe diameter.  The Dieterich-Standard ``Annubar'' is an example of an averaging Pitot tube element.

\vskip 10pt

{\it Paddle} or ``drag disk'' elements measure the raw force applied to an obstacle hanging in the flowstream, caused by fluid stagnation on the upstream face.

\vskip 10pt

{\it Flow tubes} are short venturi tubes.  {\it Flow nozzles} are venturi-like structures that clamp between pipe flanges kind of like an orifice plate.

\vskip 10pt

A {\it V-cone} element is kind of like an orifice plate in reverse: flow passes through an annular passage, around a ``donut hole'' in the middle of the pipe.

\vskip 10pt

{\it Segmental wedges} are nothing more than pipe segments with a wedge-shaped obstruction acting as a restriction.  They are most useful for measuring slurries.

\vskip 10pt

{\it Pipe elbows} may also be used as flow elements, albeit inaccurate ones.

\vskip 10pt

The amount of permanent pressure loss caused by a flow-sensing element may be a significant factor in the long-term cost of the device.  If such a loss constitutes an energy cost, then it may pay to invest in a more expensive low-loss element such as a venturi tube over a cheap orifice plate.








\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} {\bf In what ways may any of these flowmeter types be ``fooled'' to report a false flow measurement?}
\item{} Is square-root characterization necessary for all of these flow elements, or only for some of them?
\item{} Pitot tube elements are notoriously sensitive to positioning in the pipe.  For which type of flow regime will this positioning be most sensitive: {\it laminar} or {\it turbulent}?
\item{} Why might one choose to install an {\it averaging} Pitot tube instead of a classic Pitot tube?
\item{} Identify a practical application for a {\it segmental wedge} flow element.
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Continuous Fluid Flow Measurement (other differential producers)

%(END_NOTES)


