
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Law of Continuity'' subsection of the ``Fluid Mechanics'' section of the ``Physics'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.


\underbar{file i04032}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

The Law of Mass Conservation dictates that the mass flow rate through a pipe having no leaks or pulsations in the flow stream must remain constant, like current in a series circuit.  This mass flow rate is equal to the mass density of the fluid ($\rho$) multiplied by the cross-sectional area of the pipe ($A$) multiplied by the average velocity of the fluid ($\overline{v}$):

$$W = \rho A \overline{v}$$

In situations where the fluid is ``incompressible'' (i.e. does not exhibit substantial changes in density), the $\rho$ term may be ignored, and constant volumetric flow ($Q$) assumed at all places in a pipe:

$$Q = A \overline{v}$$

This tells us that the velocity of a fluid will be inversely proportional to the cross-sectional area of the pipe: if the pipe gets skinnier, the velocity increases.







\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain how the Law of Continuity for fluids is a direct consequence of the {\it Law of Mass Conservation}.
\item{} Does the Law of Continuity still hold true if the fluid in question is compressible?
\item{} Explain why all the continuity formulae specify {\it average} velocity rather than absolute velocity. 
\item{} Suppose the Reynolds number of a fluid moving through a pipe was too low to direct measure using a particular flowmeter technology.  How can we re-size the pipe (and flowmeter) to make this flow stream measureable?
\item{} If a pipe reduces from 10" diameter to 4" diameter, what happens to the average velocity of an incompressible fluid flowing through it?  {\it Answer: velocity increases by a factor of 6.25}
\item{} If a pipe reduces from 30" diameter to 18" diameter, what happens to the average velocity of an incompressible fluid flowing through it?  {\it Answer: velocity increases by a factor of 2.78}
\item{} If a pipe expands from 2" diameter to 7" diameter, what happens to the average velocity of an incompressible fluid flowing through it?  {\it Answer: velocity decreases by a factor of 12.25}
\item{} If a pipe expands from 8" diameter to 24" diameter, what happens to the average velocity of an incompressible fluid flowing through it?  {\it Answer: velocity decreases by a factor of 9}
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Fluid Mechanics (Reynolds number)

%(END_NOTES)


