
%(BEGIN_QUESTION)
% Copyright 2015, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Refer to pages 4 and 5 of the ``Micro Motion `ELITE' Coriolis Flow and Density Meters'' product datasheet (publication PS-00374 Revision L), and answer the following questions:

\vskip 10pt

Compare the turndown performance of a Coriolis flowmeter against that of a typical orifice plate flowmeter, and identify which one has better performance.  Explain {\it why} one has better performance than the other.

\vskip 10pt

Examine the graph of accuracy versus flow rate on page 4 and explain the meaning of the ``turndown ratio'' limits shown on the graph (e.g. 100:1, 20:1, 2:1).  Explain what {\it turndown} means for any measuring instrument.

\vskip 10pt

Select an appropriate model of flowmeter for measuring the flow rate of water up to 25 GPM.

\vskip 10pt

Select an appropriate model of flowmeter for measuring the flow rate of natural gas up to 400 SCFM (at a line pressure of 500 PSI).

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Page 10 contains tables showing the effect of process temperature and process pressure on measurement accuracy, both for flow rate and for density.  Explain why changes in process pressure and/or process temperature would have this effect on a Coriolis flowmeter.  
\item{} Pages 17 through 22 show illustrations of these Coriolis flowmeters.  For each of the given drawings, identify where the two vibrating tubes are located, and what shape those tubes take.
\end{itemize}

\underbar{file i04076}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

``Turndown'' refers to the ratio of lowest flow to highest flow measurable while remaining within a stated accuracy of measurement.

\vskip 10pt

Coriolis flowmeters have {\it far} better turndown performance than orifice plate flowmeters!  The quadratic characteristic of orifice plates is the principal reason for poor turndown of that flow element.  Coriolis flowmeters, by contrast, may achieve turndown ratios of 500:1 with a flow measurement accuracy of $\pm$1.25\% (page 4).  With $\pm$0.25\% accuracy, the turndown for this model of Coriolis flowmeter is 100:1.

\vskip 10pt

Best model for 25 GPM liquid flow: {\bf CMF050} (up to 30 GPM), from table at the top of page 4.

\vskip 10pt

Best model for 400 SCFM natural gas flow: {\bf CMF050} (up to 970 SCFM), from table on page 5.  This assumes a pressure drop across the meter of about 50 PSID.  If less pressure drop is required, the model {\bf CMF100} would be a better choice.

%INDEX% Reading assignment: Rosemount Micro Motion ``ELITE'' Coriolis mass/density flow transmitter datasheet

%(END_NOTES)


