
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read selected portions of the ``Daniel Ultrasonic Gas Flowmeter'' manual for the 3400 series SeniorSonic and JuniorSonic flowmeters (part number 3-9000-740 Revision H), and answer the following questions:

\vskip 10pt

Read page 3-22 and identify the minimum straight-pipe lengths upstream and downstream required for proper operation of the flowmeter.

\vskip 10pt

Read page 3-22 and identify how closely the diameter of the meter flowtube must match the inside diameter of the pipe it connects to.

\vskip 10pt

Read page 3-22 and identify how closely a temperature-sensing probe (i.e. thermowell with RTD) may be installed to the meter flowtube, and which side (upstream or downstream) it should be on.

\vskip 10pt

Read pages 5-1 through 5-5 and identify the operating principle (Doppler or transit-time) used in both the SeniorSonic and JuniorSonic gas flowmeters.


\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Why should the Senior flowmeter be installed with chords oriented horizontally?
\item{} Why should the Junior flowmeter be installed with chords 45 degrees off vertical?
\item{} How many paths (chords) are used in the Senior versus the Junior models?
\item{} How does the general design of the Senior model differ from that of the Junior model?
\end{itemize}

\underbar{file i04070}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

10D upstream and 5D downstream for the SeniorSonic meters.  20D upstream and 5D downstream for the JuniorSonic meters (page 3-22).  If bidirectional flow is desired, both sides must meet the respective upstream requirement (10D for Senior, 20D for Junior).

\vskip 10pt

The bores must match by $\pm$ 1\%, in order to meet the AGA9 standard.

\vskip 10pt

The temperature probe must be at least 3D downstream.

\vskip 10pt

This is a transit-time flowmeter (pages 5-2 and 5-3).

\vskip 10pt

The Senior flowmeter uses four-chord sensing (page 2-5) while the Junior model uses either single-chord (page 2-6) or two-chord (page 2-7) sensing.  Also, the chord paths in the Senior model run from transducer to transducer (page 5-3), while the Junior model bounces the sound wave from transducer to pipe wall to transducer (page 5-4).

\vskip 10pt

Chord orientation is recommended horizontal for the Senior model and 45$^{o}$ off vertical for the Junior model to avoid possible interference by liquids in the pipe (either resting along the bottom of the pipe and/or in the transducer ports).

%INDEX% Reading assignment: Daniel SeniorSonic and JuniorSonic ultrasonic gas flowmeter manual

%(END_NOTES)


