%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Electric Actuators'' subsection of the ``Control Valve Actuators'' section of the ``Control Valves'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\vskip 10pt

If you would like to learn more about the various types of gear mechanisms used inside electric valve actuators to reduce the electric motor's speed and boost torque necessary to move a large valve, feel free to refer to the ``Simple Machines'' section of the ``Physics'' chapter in your {\it Lessons In Industrial Instrumentation} textbook. 

\underbar{file i04184}
%(END_QUESTION)




%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Electric actuators use an electric motor to open and close the valve mechanism.  In the past, electric actuators were only used for on/off ``shutoff'' valve applications, but are now used for throttling applications as well when equipped with electronic positioners.  Gear reduction mechanisms are used inside of electric actuators to reduce motor speed (and boost torque) to move the valve.  The ``worm gear'' type of mechanism is popular: a screw turned by the high-speed motor engages with teeth on a normal gear to form a large speed-reduction ratio.

\vskip 10pt

Hand wheels and clutches often part of electric actuators.  Hand wheel provdes a way to move the valve if power fails, and the clutch prevents the wheel from spinning when the motor does.

\vskip 10pt

Some electric actuators may run off of battery packs in event of power failure.  Actuators require feedback for valve position, either switches or sensor.












\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify some advantages electric actuating systems enjoy over pneumatic actuators.
\item{} Identify some disadvantages electric actuating systems exhibit compared to pneumatic actuators.
\item{} Explain how a ``worm gear'' mechanism works (refer to the ``Simple Machines'' section of the ``Physics'' chapter in the textbook if necessary).
\item{} Identify the location of the worm gear components in the photograph showing the blue MOV.
\item{} {\bf Identify the location of the worm gear components in the photograph showing the grey MOV}.
\item{} Explain why electric valve actuators are equipped with clutch mechanisms.
\end{itemize}




%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Control Valves (globe valves)

%(END_NOTES)


