%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the introduction to the ``Stoichiometry'' section of the ``Chemistry'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i04106}
%(END_QUESTION)




%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Stoichiometry is the act of ``balancing'' the compound multipliers in a chemical equation so that the numbers of each element are equal on both sides.  In other words, it is an accounting of elemental mass, based on the Conservation of Mass.









\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain what the {\it Conservation of Mass} has to do with stoichiometry.
\item{} Examining the reaction formula showing the combustion of methane, how many moles of oxygen will be required to fully combust 15 moles of methane?
\item{} Examining the reaction formula showing the combustion of methane, how many moles of oxygen will be required to fully combust 40 moles of methane?
\end{itemize}









\vfil \eject

\noindent
{\bf Prep Quiz}

Determine whether or not the following chemical equation is {\it balanced}, and explain why it is or is not (being as specific as you can in your answer!):

$$\hbox{C}_2 + \hbox{H}_2\hbox{O} \to \hbox{H}_2 + \hbox{CO}$$








\vfil \eject

\noindent
{\bf Prep Quiz}

Determine whether or not the following chemical equation is {\it balanced}, and explain why it is or is not (being as specific as you can in your answer!):

$$2\hbox{H}_2\hbox{O}_2 \to 2\hbox{H}_2\hbox{O} + \hbox{O}_2$$









\vfil \eject

\noindent
{\bf Prep Quiz}

Determine whether or not the following chemical equation is {\it balanced}, and explain why it is or is not (being as specific as you can in your answer!):

$$\hbox{CH}_4 + \hbox{H}_2\hbox{O} \to 2\hbox{H}_2 + \hbox{CO}_2$$











\vfil \eject

\noindent
{\bf Summary Quiz}

Write a balanced chemical equation showing the complete combustion of acetylene gas (C$_{2}$H$_{2}$) with oxygen gas (O$_{2}$), resulting in water vapor (H$_{2}$O) and carbon dioxide (CO$_{2}$):

\vfil \eject

{\it Answer to summary quiz:}

$$2\hbox{C}_2\hbox{H}_2 + 5\hbox{O}_2 \to 2\hbox{H}_2\hbox{O} + 4\hbox{CO}_2$$



%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Chemistry (stoichiometry)

%(END_NOTES)


