
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read selected portions of the Fisher ``E-Body'' control valve product flier (document PF51.1:E) and answer the following questions:

\vskip 10pt

Identify the trim design of the ``Easy-E'' series: is it a stem-guided, port-guided, or cage-guided design?  Note: there is an exception called the type ``EZ'' trim!

\vskip 10pt

Page 6 of this flier has a table listing different applications and specialty trim types for the E-body valve design.  Identify some of these applications and trim styles.

\vskip 10pt

Page 7 contains an illustration of a {\it reverse-acting} globe valve body.  Examine this illustration and explain how the operation of this valve body differs from that of a ``direct-acting'' valve.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} The ``Standard Shutoff Class'' rating shown in the far-right column of the table on page 6 describes how tightly the valve trim shuts off flow when in the fully-closed position, with a rating of I (Roman numeral 1) being the leakiest and VI (Roman numeral 6) being the tightest.  Based on this table, what factor(s) contribute to a globe valve's ability to tightly shut off?
\item{} Explain how the reverse-acting valve shown on page 7 works.
\item{} Identify the principle design difference between the ED and ET trim types shown in this document.
\end{itemize}

\underbar{file i04186}
%(END_QUESTION)




%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

Cage-guided globe valve design (page 2).  The ``EZ'' trim is a stem-guided (``post-guided'' on page 6) globe valve (page 7).

\vskip 10pt

\begin{itemize}
\item{} Stringent shutoff (T trim) requires a ``soft seat'' and elastomer cage seal
\item{} General service (D trim) has metal seat and graphite cage seal
\item{} Temp to 538 deg C (S trim) has metal or ``soft'' seat and unbalanced plug (no cage seal)
\item{} Viscous or non-lubricating (Z trim) has metal or "soft" seat and post-guided plug (no cage)
\end{itemize}

\vskip 10pt

A reverse-acting valve is shown on page 7.  The ports open up further as the stem is pushed into the body.





%INDEX% Reading assignment: Fisher Easy-E body control valve product flier

%(END_NOTES)


