
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Direct/reverse actions'' subsection of the ``Valve Failure Mode'' section of the ``Control Valves'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i04192}
%(END_QUESTION)




%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

For pneumatically-actuated control valves, both the valve body and the pneumatic actuator are independently classified as either {\it direct-acting} or {\it reverse-acting}.  Thus, there is no such thing as a direct-acting valve or a reverse-acting valve, but rather we have combinations of direct- and/or reverse-acting bodies and actuators.  It is the particular combination of actuator and body action that makes a valve behave as air-to-open or air-to-close.

\vskip 10pt

DIRECT ACTUATOR: stem pushes down with air pressure

REVERSE ACTUATOR: stem pulls up with air pressure

\vskip 10pt

DIRECT VALVE BODY: opens up as stem lifts

REVERSE VALVE BODY: shuts off as stem lifts

\vskip 10pt

Direct valve body + reverse actuator = air-to-open (fail closed)

Direct valve body + direct actuator = air-to-close (fail open)

\vskip 10pt

Reverse valve body + direct actuator = air-to-close (fail open)

Reverse valve body + direct actuator = air-to-open (fail closed)

\vskip 10pt

Direct-acting valve bodies by far the most popular.




\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Why do you suppose different failure modes exist for control valves.  Why, for example, aren't all control valves made to be air-to-open (fail-closed)?
\item{} Which actuator/valve actions do you have in lab?
\end{itemize}




%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Control Valves (direct/reverse actions)

%(END_NOTES)


