
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Importance of Proper Valve Sizing'' subsection of the ``Control Valve Sizing'' section of the ``Control Valves'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i04226}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

$C_v$ is defined as the maximum GPM flow rate of water that can pass through a wide-open control valve at a pressure drop of 1 PSID.  Obviously, control valves need to be sized large enough to pass the maximum desired amount of process fluid flow.  However, an over-sized control valve is overly sensitive to small stem motions, and makes fine control of flow difficult if not impossible.  Over-sized valves are common in industry, as valves may be purchased with greater future flow rates in mind.

\vskip 10pt

A solution to the future expansion dilemma is to purchase {\it reduced-port} trim for the control valve, so that when the day comes that you need more flow capacity, you may simply change out the trim to regular (full-port) size without having to replace the entire control valve.












\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify the danger(s) of under-sizing a control valve.
\item{} Identify the danger(s) of over-sizing a control valve.
\item{} Explain why control valve over-sizing is a common problem in industry.
\item{} Explain why anyone would intentionally install a control valve with {\it reduced-port} trim.
\end{itemize}












\vfil \eject

\noindent
{\bf Prep Quiz:}

Define $C_v$ for a control valve.  Be as detailed as you can in your answer.







\vfil \eject

\noindent
{\bf Prep Quiz:}

Identify one problem that may result from the installation of an {\it over-sized} control valve.



%INDEX% Reading assignment: Lessons In Industrial Instrumentation, control valves (importance of proper sizing)

%(END_NOTES)


