
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Examine the fluid power diagram on page 2 of the Bettis ``DeltaMatic Rate-of-Drop Linebreak Detection System'' bulletin (document 41.00 Revision 10/03) and answer the following questions:

\vskip 10pt

The Differential Pressure Pilot (part \#120) signals the line valve to actuate when the pipeline pressure suddenly decreases.  The key to this pilot valve's operation are the restrictor (part \#105) and the Rate Tank (part \#108).  Explain how this system works to detect a sudden line pressure drop, performing a ``thought experiment'' where the pipeline has been holding a constant gas pressure for a long time and then that pressure suddenly falls to nothing (assuming a nearby break in the line).

\vskip 10pt

Explain how the Differential Pressure Pilot, once ``tripped'' by the detection of sudden line pressure drop, subsequently causes the line valve to shut off (close).

\vskip 10pt

Identify whether or not this automatic shutoff system has the capability to re-open the line valve after pressure is restored, and explain your answer based on the fluid power diagram.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain why a {\it rate} of pressure drop detection scheme might be a better choice for an automatic pipeline shutoff valve to have than a simple low-pressure or high-pressure trip characteristic.
\end{itemize}

\underbar{file i04204}
%(END_QUESTION)




%(BEGIN_ANSWER)

If pressure suddenly falls, the Rate Tank maintains gas pressure to the left-hand side of the Differential Pressure Pilot valve, while the (lesser) instantaneous line gas pressure is applied to the right-hand side.  This causes the Pilot to actuate.  When the left-hand and right-hand actuating pressures to the Pilot are equal, the Pilot's spring maintains it in the ``normal'' position (shown in the diagram).

\vskip 10pt

Since the shutdown system signal goes to the ``Close'' 3-way valve only, and not to the ``Open'' 3-way valve, this system does {\it not} have the capability of automatically re-opening the gas pipeline valve.

\vskip 10pt

When the Differential Pressure Pilot valve (part \#120) actuates due to a sudden drop in line pressure, it bleeds signal pressure from the actuator of the Reversing Relay (part \#25).  When this Relay goes to its ``normal'' position, it passes signal pressure to actuator of the 3-way ``Close'' valve (part \#5), sending signal pressure to the left-hand Gas/Hydraulic Tank (part \#3), sending oil to the left-hand side of the Operator (part \#1).

%(END_ANSWER)





%(BEGIN_NOTES)


%INDEX% Reading assignment: Bettis ``DeltaMatic'' datasheet

%(END_NOTES)


