
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

\vbox{\hrule \hbox{\strut \vrule{} {\bf Desktop Process exercise} \vrule} \hrule}

\noindent
Configure the controller as follows (for ``proportional-only'' control):

\begin{itemize}
\item{} Control action = {\it reverse}
\item{} Gain = {\it one-quarter the value yielding optimal control} = {\it four times the proportional band value yielding optimal control}
\item{} Reset (Integral) = {\it minimum effect} = {\it 100+ minutes/repeat} = {\it 0 repeats/minute}
\item{} Rate (Derivative) = {\it minimum effect} = {\it 0 minutes} 
\end{itemize}

In this exercise you will purposely establish a large proportional-only offset, then introduce reset (integral) action to correct this offset.  I recommend switching the controller to manual mode, adjusting the output until a low value for process variable (PV) is reached, then switching back to automatic mode and adjusting the setpoint (SP) toward the top end of the scale.  The PV should exhibit offset (inability to equal SP) due to proportional action.

\vskip 10pt

Once the controller is exhibiting a significant amount of offset, adjust the controller's reset (integral) setting to a value of 1 repeat per minute (equal to 1 minute per repeat).  Observe the controller's output value carefully as integral action begins to compensate for the offset.  Once the controller has finally eliminated offset between PV and SP, feel free to change the setpoint value to create another offset and then observe integral action at work once again.

Explain in your own words what the controller's integral action is doing to eliminate the ``offset'' that proportional action can never completely eliminate.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} If you desired a swifter integral response from the loop controller, what could you alter in this system to make the integral action respond more aggressively?
\end{itemize}

\underbar{file i04289}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

{\bf Lesson:} watching how integral (reset) action gets rid of proportional-only offset over time.

%INDEX% Desktop Process: automatic control of motor speed (demonstration of integral action)

%(END_NOTES)


