
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Automatic and Manual Modes'' subsection of the ``Pneumatic PID Controllers'' section of the ``Closed-Loop Control'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i04265}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

A transfer valve switches between automatic mode (output pressure caused by the balance mechanism) and manual mode (output pressure set by a regulator).  Transitioning ``bumplessly'' between automatic and manual modes requires care by the operator to make adjustments so that the two pressures are equal when the transfer valve is switched.

\vskip 10pt

The Foxboto model 43AP controller uses a curved plastic tube with a ball in the center to indicate when the two pressures (auto and manual) are balanced prior to switching.






\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain the purpose of the bias spring in the pneumatic controller mechanism shown.
\item{} Why do we care about bumpless transfer between automatic and manual modes?
\item{} Explain what would happen if the bias spring broke in half.
\item{} Suppose the controller is in automatic mode and the balance indicator ball is all the way to the right.  Describe the steps necessary to switch this controller to manual mode ``bumplessly''.
\item{} Suppose the controller is in manual mode and the balance indicator ball is all the way to the right.  Describe the steps necessary to switch this controller to automatic mode ``bumplessly''.
\end{itemize}


%INDEX% Reading assignment: Lessons In Industrial Instrumentation, closed-loop control (pneumatic PID controllers, auto and manual modes)

%(END_NOTES)


