
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Valve Packing'' section of the ``Control Valves'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i04206}
%(END_QUESTION)




%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

``Packing'' is a leak-tight seal between the moving stem and the stationary valve body, like the stuffing box on a boat.  Fugitive emissions is when undesired leaks of process fluid occur (as opposed to normal emissions out of stacks).  Packing usually takes form of rings compressed around valve stem (Teflon or graphite).  Sometimes a ``lantern ring'' spacer is inserted to allow lubricant to reach inside packing rings.

\vskip 10pt

If packing is under-tightened, it will leak process fluid.  If packing is over-tightened, it will bind the valve stem's motion and wear at an accelerated rate.  The proper remedy for a leaking packing (if properly torqued) is replacement, not further tightening.

\vskip 10pt

If packing itself provides the elastic resistance, it is called ``stationary loading.''  If a metal spring is included in the packing stack to help maintain proper tension, it is called ``live loading.''  This loading spring may be internal with the packing rings, or be a set of Belleville washers between the packing flange and the packing follower.

\vskip 10pt

\noindent
Packing lubricators act as grease pumps to force lubricant into packing assembly.  Here is a typical operating procedure:

\item{} Shut block valve
\item{} Unthread bolt
\item{} Insert grease into lubricator body
\item{} Thread bolt into hole to generate grease pressure
\item{} Open block valve and thread bolt fully into hole
\item{} Shut block valve
\end{itemize}

\vskip 10pt

Teflon packing is 5 to 10 times less friction than graphite packing, but cannot tolerate hard nuclear radiation or high temperatures.  Graphite packing allows galvanic corrosion to occur between bonnet and stem metals.  Graphite/Teflon hybrids exist.  Asbestos used to be the packing material of choice, before it was classified as a hazardous material.

\vskip 10pt

A ``bellows'' seal is an accordian-shaped metal tube fixed to the valve stem on one end and fixed to the valve body on the other, allowing the valve stem to move while presenting absolutely no leak path for process fluid to escape.  However, since bellows can rupture, the valve is still equipped with a conventional packing to prevent catastrophic fluid leakage.  A pressure sensor connected between the bellows seal and the conventional packing may be used to detect such a rupture.




\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify some of the various packing materials available for control valves, and their application limits.
\item{} Describe the function of ``live-loaded'' control valve packing and what the alternative is.
\item{} Explain why bellows packing must always be accompanied by conventional friction-packing rather than be used alone.
\item{} Explain how you could calculate the amount of friction offered by valve packing, just by testing the valve's motion using an adjustable air pressure regulator and a pressure gauge.
\end{itemize}







\vfil \eject

\noindent
{\bf Summary Quiz:}

Applications where graphite valve packing would be superior to Teflon valve packing include:

\begin{itemize}
\item{} Where the process fluid temperature is extremely hot
\vskip 5pt 
\item{} Where fugitive emissions of the process fluid are a problem
\vskip 5pt 
\item{} Where stem friction must be absolutely minimized
\vskip 5pt 
\item{} Where galvanic corrosion of the stem is a problem
\vskip 5pt 
\item{} Where the packing must operate without grease lubrication
\vskip 5pt 
\item{} Where materials cost must be held to a minimum
\end{itemize}



%INDEX% Reading assignment: Lessons In Industrial Instrumentation, Control Valves (packing)

%(END_NOTES)


