
%(BEGIN_QUESTION)
% Copyright 2009, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

\vbox{\hrule \hbox{\strut \vrule{} {\bf Desktop Process exercise} \vrule} \hrule}

\noindent
Configure the controller as follows (for ``proportional-only'' control):

\begin{itemize}
\item{} Control action = {\it reverse}
\item{} Gain = {\it set to whatever value yields optimal control}
\item{} Reset (Integral) = {\it minimum effect} = {\it 100+ minutes/repeat} = {\it 0 repeats/minute}
\item{} Rate (Derivative) = {\it minimum effect} = {\it 0 minutes} 
\end{itemize}

You will now purposely create a condition of proportional-only offset.  I recommend switching the controller to manual mode, adjusting the output until a low value for process variable (PV) is reached, then switching back to automatic mode and adjusting the setpoint (SP) toward the top end of the scale.  The PV should exhibit offset (inability to equal SP) due to proportional action.

Once the controller is exhibiting a significant amount of offset, practice using the controller's ``manual'' mode to eliminate the offset.  For example, if the SP is 70\% and the PV is 63\% in automatic mode, switch to manual mode and manually adjust the output to get the PV equal to 70\% (the desired value) then switch back to automatic mode.

After eliminating offset, move the SP to some new value to re-create an offset again and practice using manual mode to eliminate the offset once more.

\vskip 10pt

Explain how the common feature of {\it setpoint tracking} programmed into the controller helps you in your task of manually eliminating offset.

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Explain how the feature of {\it output tracking} is also seen at work in your controller during this exercise.
\end{itemize}

\underbar{file i04286}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

{\bf Lesson:} using manual mode in a controller with setpoint and output tracking to reset proportional-only offset.


%INDEX% Desktop Process: automatic control of motor speed (manually re-setting P-only offset)

%(END_NOTES)


