
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Read and outline the ``Motor Contactors'' subsection of the ``On/Off Electric Motor Control Circuits'' section of the ``Discrete Control Elements'' chapter in your {\it Lessons In Industrial Instrumentation} textbook.  Note the page numbers where important illustrations, photographs, equations, tables, and other relevant details are found.  Prepare to thoughtfully discuss with your instructor and classmates the concepts and examples explored in this reading.

\underbar{file i04495}
%(END_QUESTION)





%(BEGIN_ANSWER)


%(END_ANSWER)





%(BEGIN_NOTES)

A {\it contactor} is nothing more than a three-pole, single-throw (3PST) relay with normally-open contacts used to interrupt power to a motor.  Contactor coils are typically rated for 120 VAC, while contacts might be 480 VAC or greater.

\vskip 10pt

Arc shield covers protect personnel from the hazards of electric shock as well as arc flash and arc blast when the contacts interrupt high currents to the motor.











\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Identify the safety hazard(s) associated with actuating a motor contactor with the arc shields removed.
\item{} Define {\it arc flash}, and describe how this could potentially occur within a motor contactor (especially if a portion of the contactor happened to be damaged).
\end{itemize}











\vfil \eject

\noindent
{\bf Prep Quiz:}

A {\it contactor} is:

\begin{itemize}
\item{} Two metal blocks used to interrupt electric current in a circuit
\vskip 5pt 
\item{} A special overcurrent-protective device used on electric motors
\vskip 5pt 
\item{} A tool used to check for close mechanical contact on a motor shaft
\vskip 5pt 
\item{} A device used to maintain proper wire tension on a terminal strip
\vskip 5pt 
\item{} A heavy-duty relay used to switch power to an electric motor
\vskip 5pt 
\item{} A special overvoltage-protective device used on electric motors
\end{itemize}

%INDEX% Reading assignment: Lessons In Industrial Instrumentation, motor contactors

%(END_NOTES)

