
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Suppose we have an H1 Fieldbus segment with 6 devices connected.  The LAS, of course, maintains these addresses in its ``Live List'' so it knows who to send tokens to:

\begin{itemize}
\item{} Address 24: Multi-channel temperature transmitter
\item{} Address 33: Control valve positioner
\item{} Address 35: Control valve positioner
\item{} Address 40: Pressure transmitter
\item{} Address 88: Level transmitter
\item{} Address 107: Local indicator
\end{itemize}

If someone were to connect a new device with an address of 249, how would the LAS ever know it was there, and how would it get added to the Live List?  Conversely, how does a device fall off the Live List?

\vskip 20pt \vbox{\hrule \hbox{\strut \vrule{} {\bf Suggestions for Socratic discussion} \vrule} \hrule}

\begin{itemize}
\item{} Suppose someone were to plug a new device into one of the coupler ports on this H1 segment, and this new device happened to have a pre-assigned address of 40.  What would happen as a result?
\item{} Suppose someone were to plug a new device into one of the coupler ports on this H1 segment, and this new device happened to have a pre-assigned address falling within the FUN/NUN address range.  What would happen as a result?
\end{itemize}

\underbar{file i04571}
%(END_QUESTION)





%(BEGIN_ANSWER)

Hint: the PN (Probe Node) token!

%(END_ANSWER)





%(BEGIN_NOTES)

The LAS periodically issues ``Probe Node'' tokens during unscheduled communication times to probe for new nodes.  A new device connected to the segment -- when it receives a PN with its address attached -- will respond to the LAS and be added to the Live List.  If a device ever repeatedly fails to respond to specific tokens issued by the LAS, it will be removed from the Live List.

The LAS also periodically transmits the latest Live List to any backup Link Master devices in case they must take over as the primary LAS.

%INDEX% Fieldbus, FOUNDATION (H1): Device addresses
%INDEX% Fieldbus, FOUNDATION (H1): Probe Node (PN) token

%(END_NOTES)


