
%(BEGIN_QUESTION)
% Copyright 2010, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

Write the appropriate letter next to each network arbitration method, from the definitions given below:

\vskip 10pt

\begin{itemize}
\item{} Master/Slave: \underbar{\hskip 50pt}
\vskip 10pt
\item{} Token: \underbar{\hskip 50pt}
\vskip 10pt
\item{} CSMA/CD: \underbar{\hskip 50pt}
\vskip 10pt
\item{} CSMA/BA: \underbar{\hskip 50pt}
\vskip 10pt
\item{} CSMA/CA: \underbar{\hskip 50pt}
\end{itemize}

\vskip 20pt

\begin{itemize}
\item{(A)} Any device may initiate communication at any time, leading to inevitable ``collisions.'' Collisions are detected, and then the colliding devices speak again after a time delay set by a unique, pre-programmed ``back-off'' time.
\item{(B)} Only select devices get to initiate communication.  All others are merely allowed to respond to queries.
\item{(C)} Devices each have a unique, pre-programmed wait time which must elapse after detecting a silent network before they may transmit.
\item{(D)} Each device gets a time to initiate communication, temporary mastership being designated by a special data message passed from one device to the next.
\item{(E)} Any device may initiate communication at any time, leading to inevitable ``collisions.''  Collisions are detected, and then the colliding devices each try speaking again after a random ``back-off'' time.
\item{(F)} Each device is reserved a fixed time slot in a repeating schedule in which to initiate communication.
\end{itemize}

\underbar{file i04561}
%(END_QUESTION)





%(BEGIN_ANSWER)

\begin{itemize}
\item{} Master/Slave: \underbar{\bf B}
\vskip 10pt
\item{} Token: \underbar{\bf D}
\vskip 10pt
\item{} CSMA/CD: \underbar{\bf E}
\vskip 10pt
\item{} CSMA/BA: \underbar{\bf A}
\vskip 10pt
\item{} CSMA/CA: \underbar{\bf C}
\end{itemize}

%(END_ANSWER)





%(BEGIN_NOTES)

{\bf This question is intended for exams only and not worksheets!}.

%(END_NOTES)


