
 Copyright 2014, Tony R. Kuphaldt, released under the Creative Commons Attribution License (v 1.0)
% This means you may do almost anything with this work of mine, so long as you give me proper credit

%(BEGIN_FRONTMATTER)

\centerline{\bf General tool and supply list}

\vskip 10pt

\noindent
{\bf Wrenches}

\item{$\bullet$} Combination (box- and open-end) wrench set, 1/4" to 3/4" -- {\it the most important wrench sizes are 7/16", 1/2", 9/16", and 5/8"; get these immediately!}
\item{$\bullet$} Adjustable wrench, 6" handle (sometimes called ``Crescent'' wrench)
\item{$\bullet$} Hex wrench (``Allen'' wrench) set, fractional -- 1/16" to 3/8"
\item{$\bullet$} {\it Optional:} Hex wrench (``Allen'' wrench) set, metric -- 1.5 mm to 10 mm
\item{$\bullet$} {\it Optional:} Miniature combination wrench set, 3/32" to 1/4" (sometimes called an ``ignition wrench'' set)

\vskip 5pt

Note: {\it when turning any threaded fastener, one should choose a tool engaging the maximum amount of surface area on the fastener's head in order to reduce stress on that fastener.  (e.g. Using box-end wrenches instead of adjustable wrenches; using the proper size and type of screwdriver; never using any tool that mars the fastener such as pliers or vise-grips unless absolutely necessary.)}

\vskip 10pt

\noindent
{\bf Pliers}

\item{$\bullet$} Needle-nose pliers
\item{$\bullet$} Tongue-and-groove pliers (sometimes called ``Channel-lock'' pliers)
\item{$\bullet$} Diagonal wire cutters (sometimes called ``dikes'')
\medskip

\vskip 10pt

\noindent
{\bf Screwdrivers}

\item{$\bullet$} Slotted, 1/8" and 1/4" shaft
\item{$\bullet$} Phillips, \#1 and \#2
\item{$\bullet$} Jeweler's screwdriver set
\item{$\bullet$} {\it Optional:} Magnetic multi-bit screwdriver (e.g. Klein Tools model 70035)

\vskip 10pt

\noindent
{\bf Electrical}

\item{$\bullet$} Multimeter, Fluke model 87-IV or better
\item{$\bullet$} Alligator-clip jumper wires
\item{$\bullet$} Soldering iron (10 to 40 watt) and rosin-core solder
\item{$\bullet$} Resistor, potentiometer, diode assortments (from first-year lab kits)
\item{$\bullet$} Package of insulated compression-style fork terminals (14 to 18 AWG wire size, \#10 stud size)
\item{$\bullet$} Wire strippers/terminal crimpers for 10 AWG to 18 AWG wire and insulated terminals
\item{$\bullet$} {\it Optional:} ratcheting terminal crimp tool (e.g. Paladin 1305, Ferrules Direct FDT10011, or equivalent)
\medskip

\vskip 10pt

\noindent
{\bf Safety}

\item{$\bullet$} Safety glasses or goggles (available at BTC bookstore)
\item{$\bullet$} Earplugs (available at BTC bookstore)
\medskip

\vskip 10pt

\noindent
{\bf Miscellaneous}

\item{$\bullet$} Simple scientific calculator (non-programmable, non-graphing, no conversions), TI-30Xa or TI-30XIIS recommended.  Required for some exams!  %INSTRUCTOR {\bf Demonstrate to each and every student how to use the memory functions (STO, RCL) on their calculators!}
\item{$\bullet$} Portable personal computer with Ethernet port and wireless.  Windows OS strongly preferred, tablets discouraged.
\item{$\bullet$} Masking tape (for making temporary labels)
\item{$\bullet$} Permanent marker pen
\item{$\bullet$} Teflon pipe tape
\item{$\bullet$} Utility knife
\item{$\bullet$} Tape measure, 12 feet minimum
\item{$\bullet$} Flashlight
\medskip

\vskip 10pt

An inexpensive source of tools is your local pawn shop.  Look for tools with unlimited lifetime guarantees (e.g. {\it Sears} ``Craftsman'' brand).  Check for BTC student discounts as well!



\vfil

\underbar{file {\tt tools}}
\eject
%(END_FRONTMATTER)


