
\chapter{Contributors}

This is an open-source book, which means everyone has a legal write to modify it to their liking.  As the author, I freely accept input from readers that will make this book better.  This appendix exists to give credit to those readers who have made substantial contributions to this book.

Sadly, this list does not show the names of \textit{every} person who has helped me identify and correct minor typographical and grammatical errors.  The list of names and errors would be quite substantial, I must admit.  Those persons who are listed for their identification of typographical errors have earned a place on the list through sheer volume of errors found.  I am indebted to my students, and to readers around the world for their careful reading of the text and their considerate feedback.

\vskip 10pt

If you find this textbook exceeds your expectations, know that it is primarily because of this detailed reader feedback.  When a book is edited by its author, based on daily feedback from the people using it, errors as well as unclear explanations get identified and corrected at a much more rapid pace than what is typical during a traditional (published) textbook editing cycle.  As an author, you might think you have explained something perfectly well, only to find your readers mystified.  It is only by discovering your readers are baffled by your ``good'' explanation that you will know to edit that explanation, or to take a completely different approach in explaining the concept.  Then, you need to get more feedback after the edit(s), to see whether or not you've made an improvement.  

This kind of tight feedback loop between authors and readers simply doesn't happen in a traditional textbook publishing model.  In an open-source, self-published model, however, this kind of feedback is not only possible, but is also typical.  For this reason, I strongly encourage teachers everywhere to consider writing their own open-source textbooks, using their students as editors to prove whether or not their writing is effective.  The world is unfortunately filled with poor-quality texts.  Let's change this sad state of affairs, one open-source textbook at a time!





\filbreak
\section{Error corrections}

\vskip 10pt

\noindent
\textbf{Abdelaziz, Ahmed (June 2014)} 

\begin{itemize}
\item Identified numerous typographical errors throughout the book, mostly \textit{repeated repeated} words.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Ames, Ben (January 2015)} 

\begin{itemize}
\item Identified typographical error in the \textit{Problem-Solving and Diagnostic Strategies} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Balbaa, Amr (May 2015)} 

\begin{itemize}
\item Identified mathematical error in the \textit{Continuous Fluid Flow Measurement} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Barton, Bruce (January 2015, March 2015, April 2015)} 

\begin{itemize}
\item Identified typographical errors in the \textit{Discrete Control Elements} chapter regarding solenoid valve construction, in the \textit{Process Dynamics and PID Controller Tuning} chapter, and also in the \textit{Digital Data Acquisition and Networks} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Brierly, Ryan (April 2016)} 

\begin{itemize}
\item Identified typographical error in the \textit{Physics} chapter regarding transition points between laminar and turbulent flow regimes.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Brown, Kevin (February 2011)} 

\begin{itemize}
\item Identified typographical error in the \textit{Control Valves} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Bultsma, Brent (September 2011-June 2012)} 

\begin{itemize}
\item Identified multiple typographical errors throughout the book.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Chong, Yi-Vonne (August 2017)} 

\begin{itemize}
\item Identified typographical error in a protective relay wiring diagram.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Douglas, Jon (Spring 2017)} 

\begin{itemize}
\item Suggested the inclusion of \textit{self-purged impulse lines} to the book and provided documentation of this practice from industry.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Edwards, William (February 2013)} 

\begin{itemize}
\item Identified error in the \textit{Calculus} chapter, where $\Delta t$ was confused with $\Delta m$.
\end{itemize}



\vskip 10pt

\noindent
\textbf{Esher, Cynthia A. (December 2009)} 

\begin{itemize}
\item Identified error of referring to ``SAMA'' diagrams.  I changed these references to ``functional'' diagrams instead (\textit{Instrumentation Documents} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Floyd, Wade (November 2015)} 

\begin{itemize}
\item Identified an error the distance between Earth and the sun \textit{Chemistry} chapter: the distance is 93 \textit{million} miles, not 93 \textit{billion} miles as I originally had it.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Gassman, Jessica (February 2013)} 

\begin{itemize}
\item Suggested clarification in the \textit{Closed-Loop Control} chapter regarding the auto/manual transfer of pneumatic controllers.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Glundberg, Blake (February 2011, April 2011)} 

\begin{itemize}
\item Identified typographical errors (\textit{Control Valves} chapter, \textit{Digital Data Acquisition and Networks} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Ireland, Chandler (February 2015)} 

\begin{itemize}
\item Identified typographical error (\textit{Closed-Loop Control} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Joe, Savio (February 2013)} 

\begin{itemize}
\item Identified typographical error (\textit{Instrument Connections} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Ketteridge, Tim and Dieckman, Lloyd (April 2013)} 

\begin{itemize}
\item Identified conceptual errors in the IP and TCP sections of the \textit{Digital Data Acquisition and Network} chapter.  The task of portioning large data blocks into segments and packets is the job of TCP (or Application-layer protocols), not IP.  IP does, however, support fragmentation of large packets into smaller packets and reassembly into large packets again if needed for communication along network types that cannot support large packet sizes.  I had these concepts conflated.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Ketteridge, Tim (December 2013)} 

\begin{itemize}
\item Identified an error in measurement units for the Ideal Gas Law ($PV = NkT$).  When using this form of the Ideal Gas Law, $P$ needs to be in units of Pascals (not atmospheres), and $V$ needs to be in units of cubic meters (not liters).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Marques, Tiago (July 2012)} 

\begin{itemize}
\item Identified lack of origin lines in linear graphs where 4-20 mA signals are related to practical measurement ranges (\textit{Analog Electronic Instrumentation} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Morgan, Clayton (August 2018)} 

\begin{itemize}
\item Identified typographical error in the (\textit{Continuous Pressure Measurement} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Mhyre, Phil (January 2012)} 

\begin{itemize}
\item Identified spelling error on the word ``desiccant'' (\textit{Discrete Control Elements} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Nobiling, David (July 2016)} 

\begin{itemize}
\item Identified typographical error in the (\textit{Digital data acquisition and networks} chapter), where the terms ``PLC'' and ``VFD'' were interchanged.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Pulido, Alberto (May 2012)} 

\begin{itemize}
\item Identified conceptual error regarding Fieldbus terminator resistor packages (\textit{FOUNDATION Fieldbus Instrumentation} chapter).
\end{itemize}


\vskip 10pt

\noindent
\textbf{Sangani, Champa (September 2009) and Brainard, Ben (January 2012)} 

\begin{itemize}
\item Identified calculation error in milliamp-to-pH scaling problem (\textit{Analog Electronic Instrumentation} chapter).
\end{itemize}


\vskip 10pt

\noindent
\textbf{Saxton, Trevor (August 2015)} 

\begin{itemize}
\item Identified mathematical errors in the ``Phasor Arithmetic'' section of the (\textit{AC Electricity} chapter).
\end{itemize}


\vskip 10pt

\noindent
\textbf{Schultz, Steven (February 2011)} 

\begin{itemize}
\item Identified error in a pilot-loaded pressure regulator design depicted in the \textit{Control Valves} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Sloan, Wyatt (October 2012)} 

\begin{itemize}
\item Identified typographical error in the \textit{Continuous Temperature Measurement} chapter.  Later (January 2013) identified multiple places where the word ``identify'' was used instead of ``identity.''
\end{itemize}

\vskip 10pt

\noindent
\textbf{Sobczak, Jacek (October 2015)} 

\begin{itemize}
\item Identified missing files in the .tar archive, suggested changes to the \LaTeX{} source code producing better page numbering in the book's table of contents, and identified broken links on the website.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Thompson, Brice (June 2009)} 

\begin{itemize}
\item Identified errors in high/low select and high/low limit function illustrations (\textit{Basic Process Control Strategies} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{TooToonchy, Hossein (June 2014)} 

\begin{itemize}
\item Identified typographical errors in the (\textit{FOUNDATION Fieldbus Instrumentation} chapter).
\end{itemize}

\vskip 10pt

\noindent
\textbf{Tsiporenko, Michael (June 2009)} 

\begin{itemize}
\item Identified typographical error in the \textit{Introduction to Industrial Instrumentation} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Upham, Jaime (October 2018)} 

\begin{itemize}
\item Identified typographical error in the \textit{Continuous Level Measurement} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Villajulca, Jose Carlos (August 2011)} 

\begin{itemize}
\item Identified error in PID tuning response for a ``generic'' simulated process in the \textit{Process Dynamics and PID Controller Tuning} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Wingerter, John (January 2017)} 

\begin{itemize}
\item Identified an oversight in explaining safety considerations during control valve disassembly, in the \textit{Disassembly of a Sliding-Stem Control Valve} appendix.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Young, Jeanne (April-May 2012)} 

\begin{itemize}
\item Identified errors in proximity switch wire color codes and suggested correction in the \textit{Programmable Logic Controllers} chapter.
\item Identified pattern inconsistent with industry convention in motor control circuit schematic diagrams within the \textit{Discrete Control Elements} chapter.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Ziels, Matt (October 2018)} 

\begin{itemize}
\item Suggested adding lock-out/tag-out procedure to transmitter manifold valves.
\end{itemize}















\filbreak
\section{New content}

\noindent
\textbf{Coy, John J. et. al. (December 2012)} 

\begin{itemize}
\item Sampled graphic illustrations from his NASA Reference Publication ``Gearing'' (1152), showing different types of gear trains for inclusion in the \textit{Simple Machines} section of the \textit{Physics} chapter.  This NASA document, being a US government publication unclassified and approved for ``unlimited'' distribution, lies within the public domain and so may be freely sampled into other works such as this one.
\end{itemize}


\vskip 10pt

\noindent
\textbf{Dennis, Japheth (2011-2012 academic year)} 

\begin{itemize}
\item Suggested additional examples of PID controller responses to graph, to help illustrate the unique features of each action.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Faydor, L. Litvin et. al. (December 2012)} 

\begin{itemize}
\item Sampled 5-planet planetary gear set illustration from his NASA publication ``New Design and Improvement of Planetary Gear Trains'', for inclusion in the \textit{Simple Machines} section of the \textit{Physics} chapter.  This NASA document, being a US government publication unclassified and approved for ``unlimited'' distribution, lies within the public domain and so may be freely sampled into other works such as this one.  Also sampled helicopter transmission gear set illustration from his NASA publication ``Handbook on Face Gear Drives With a Spur Involute Pinion''.
\end{itemize}

\vskip 10pt

\noindent
\textbf{Goertz, Kevin (2006-2007 academic year)} 

\begin{itemize}
\item Took photographs of various flowmeters, control valves, and an insertion pH probe assembly. 
\end{itemize}

\vskip 10pt

\noindent
\textbf{Larsen, Dayn (April 2018)} 

\begin{itemize}
\item Identified typographical errors in the (\textit{Digital Data Acquisition and Networks} chapter).
\end{itemize}

\vskip 10pt


\noindent
\textbf{Marshall, Travis (2012-2013 academic year)} 

\begin{itemize}
\item Contributed analogy of a ``shadow'' to the explanation of absorption spectroscopy, after helping to install a laser stack gas analyzer during a summer internship and helping to explain that instrument's operating principle to himself.
\end{itemize}



\vskip 10pt


\noindent
\textbf{Poelma, John (2010-2011 academic year)} 

\begin{itemize}
\item Took photographs of various pressure vessels, instruments, control valves, and other process hardware at NASA's Stennis Space Center in Mississippi. 
\end{itemize}




\vskip 10pt


\noindent
\textbf{Toner, Aaron (August 2012)} 

\begin{itemize}
\item Took photographs of his self-constructed MicroLogix 1100 / C-More Micro PLC trainer for inclusion in this book. 
\end{itemize}

\vskip 10pt


























%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

